\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{An Educational Framework for AI-Driven RF Signal Processing on FPGA}

\author{\IEEEauthorblockN{Jude Eschete}
\IEEEauthorblockA{\textit{Dept. of Electrical and Computer Engineering} \\
\textit{Stevens Institute of Technology}\\
Hoboken, NJ, USA \\
JEschete@stevens.edu}
\and
\IEEEauthorblockN{Project Advisor: Bernard Yett}
\IEEEauthorblockA{\textit{Dept. of Electrical and Computer Engineering} \\
\textit{Stevens Institute of Technology}\\
Hoboken, NJ, USA \\
byett1@stevens.edu}
}

\maketitle

\begin{abstract}
University electrical engineering programs typically deliver instruction in FPGA digital design, RF communications, embedded systems, and machine learning as independent courses, leaving graduates without structured experience integrating these disciplines. This paper presents an educational framework built around a benchtop hardware platform in which STM32 microcontrollers paired with RFM95W LoRa transceivers transmit RF signals that are received and processed in real time by a Digilent Nexys A7-100T FPGA, with collected signal data feeding a machine learning classification pipeline. A five-module progressive curriculum sequences students from hardware bring-up and embedded firmware development through FPGA RTL design, system integration, and AI-driven signal analysis. Each module makes the interfaces between disciplines explicit teaching objectives, enabling students to reason about how design decisions in one domain propagate constraints into adjacent ones.
\end{abstract}

\begin{IEEEkeywords}
FPGA, LoRa, machine learning, RF signal processing, embedded systems, curriculum design, signal classification
\end{IEEEkeywords}

\section{Introduction}

Modern electrical and computer engineering practice increasingly demands fluency across multiple technical domains simultaneously. Defense, telecommunications, and aerospace systems routinely integrate real-time RF data acquisition, hardware-accelerated digital processing, embedded firmware, and intelligent signal analysis within a single pipeline. University curricula, however, frequently deliver these disciplines in isolation. A student may learn HDL synthesis and timing closure in a digital design course, modulation theory and channel models in a communications course, interrupt-driven firmware in an embedded systems course, and classification on benchmark datasets in a machine learning course—without encountering a structured experience that requires synthesizing these skills into a working system.

This isolation creates a gap between academic preparation and professional practice. Engineers who understand each domain individually often struggle with the interface problems that arise when combining them: how RF modulation parameters affect FPGA buffer sizing, how reconfigurable logic resource constraints shape a data collection pipeline for machine learning, or how real-world RF propagation deviates from classroom channel models. These cross-disciplinary concerns are not incidental details but core engineering competencies that no single-domain course addresses.

Prior research has contributed to individual components of this problem but has not combined them into an integrated educational experience. On the pedagogical side, cloud-based FPGA toolchains with structured lab sequences have lowered the barrier to digital design education \cite{b9}, and the STM32 Nucleo platform has been validated as an effective embedded systems teaching tool with significantly greater peripheral capability than Arduino-class boards \cite{b10}. On the technical side, FPGA-based RF signal processing and inference have advanced considerably: Boolean reservoir computing achieves CNN-competitive modulation classification accuracy with far fewer trainable parameters \cite{b1}, streaming CNN architectures on RFSoC FPGAs classify eight modulation schemes continuously at 128~MHz \cite{b3}, FPGA-accelerated ResNet reduced angle-of-arrival inference cycle count by 28.2\% over CPU while improving robustness to multipath fading \cite{b2}, and SoC FPGA LoRa gateways have demonstrated hardware-accelerated processing within tight resource budgets \cite{b7}. For LoRa signal classification specifically, deep learning has proven effective across multiple tasks: CNNs on spectrogram inputs achieved 99.77\% spreading factor detection across a $-$20~dB to +30~dB SNR range \cite{b5}, CNNs on raw I/Q data reached 93.3\% device identification accuracy under real experimental conditions \cite{b6}, neural network demodulation matched coherent matched-filter performance while tolerating carrier frequency offset \cite{b8}, and comparative studies of DNN, CNN, and DBN architectures provide empirical guidance on accuracy-complexity tradeoffs for resource-constrained deployment \cite{b4}. Taken together, this body of work establishes the technical feasibility of every subsystem in the proposed platform. What is absent is a curriculum that sequences these subsystems into a unified learning experience where the connections between them are the primary subject of instruction.

The contributions of this paper are: (1) a five-module progressive educational curriculum that integrates FPGA digital design, RF communications, embedded firmware, and machine learning into a single hardware-grounded learning sequence; (2) a concrete benchtop reference platform built from commercially available components that instantiates each cross-domain interface as an experimentally observable design problem; and (3) a custom software engagement simulator that provides students with a physics-grounded, labeled training environment before physical hardware is available. Section~\ref{sec:problem} formally defines the research problem and its context. Section~\ref{sec:framework} presents the framework in detail.

\section{Problem Statement}
\label{sec:problem}

The research problem is the absence of a structured, progressive educational curriculum that integrates FPGA digital design, RF communications, embedded microcontroller programming, and machine learning-based signal analysis into a unified learning experience where each cross-domain interface is a deliberate teaching objective.

The target population is senior undergraduate and master's-level electrical engineering students who have completed introductory coursework in each of the four domains individually. These students possess the foundational prerequisites---HDL syntax, basic modulation and link theory, embedded C programming, and introductory machine learning---but lack experience working with systems that require all four simultaneously. When confronted with a multidisciplinary design problem, they face a secondary difficulty: understanding how a decision made within one domain generates requirements and constraints in adjacent domains that cannot be resolved from within a single discipline.

The concrete system context is a benchtop RF signal processing testbed. Multiple STM32 Nucleo-L476RG microcontrollers, each paired with an RFM95W LoRa transceiver, serve as independently configurable beacon transmitters. A receiver station captures incoming transmissions and forwards parsed RF data to a Digilent Nexys A7-100T FPGA, which performs real-time detection, packet integrity verification, and feature extraction in programmable logic on the Artix-7 fabric. Extracted signal features are collected by a host machine and used to train and evaluate machine learning classifiers for signal classification and emitter identification tasks. This architecture exposes four concrete interface problems that span disciplinary boundaries: the firmware-to-RF interface, the RF-to-FPGA interface, the FPGA-to-data-pipeline interface, and the data-to-model interface. Each requires cross-disciplinary reasoning to characterize and resolve.

The specific problem being solved is not a shortage of technical resources in any individual domain. Comprehensive textbooks, open-source FPGA IP cores, well-documented ML frameworks, and hardware reference designs exist for every component in this system. The gap is the absence of a curriculum that sequences these resources into an integrated learning experience, identifies the interface problems between domains as first-class subjects of instruction, and provides students with a concrete hardware implementation through which to develop and test their understanding.

The problem extends to the educator as well. An instructor preparing a multidisciplinary course of this kind cannot simply draw on deep expertise in a single domain---they must understand how each discipline's design decisions affect the others well enough to write labs, anticipate failure modes, and evaluate student work at the integration points. In practice, few faculty have equivalent depth across FPGA digital design, RF communications, embedded firmware, and machine learning simultaneously; most come to one or two of these domains as relative outsiders. Without a pre-existing framework that makes the cross-domain interfaces explicit---specifying what the integration problems are, why they arise, and how they can be observed on real hardware---the course preparation burden alone may be enough to prevent such a curriculum from being developed at all. The framework proposed in this paper addresses both sides of this problem: it provides students with the integrated learning experience they currently lack, and it provides educators with the structured reference they would need to deliver it.

\section{Proposed Integrated Educational Framework}
\label{sec:framework}

The proposed solution is a five-module curriculum paired with the benchtop hardware platform described in Section~\ref{sec:problem}. The framework assumes foundational domain knowledge and targets the integration layer: the design decisions, interface protocols, and system-level reasoning skills that emerge only when students work across disciplinary boundaries simultaneously.

\subsection{Hardware Platform}

The platform was designed around three subsystem layers, with each hardware selection driven by what it asks students to think about rather than by peak performance.

The transmission layer consists of STM32 Nucleo-L476RG microcontrollers, each paired with an RFM95W LoRa transceiver operating in the 915~MHz ISM band. LoRa was chosen as the RF modality because its chirp spread spectrum waveform exposes a small set of directly tunable parameters---spreading factor, bandwidth, and coding rate---whose effects on link behavior are large enough to observe with benchtop equipment. A student who changes the spreading factor from SF7 to SF12 gains roughly 7.5~dB of link margin per step but cuts the data rate in half \cite{b12}; that single decision ripples downstream into FPGA buffer sizing, dataset collection time, and ultimately the variance of the ML training set. This chain of consequences, spanning all four disciplines, is exactly the kind of reasoning the framework is designed to develop. The STM32 platform supports this goal by providing enough peripheral flexibility---multiple independent SPI, I2C, and USART interfaces with DMA support---that students can make and measure the impact of real firmware design decisions rather than working around hardware limitations \cite{b10,b13}. Prior work confirms that the RF signatures produced by LoRa hardware are rich enough to support device identification \cite{b6} and neural network demodulation \cite{b8} at the analysis layer, validating LoRa as a suitable signal source for the full pipeline.

At the processing layer, the Digilent Nexys A7-100T serves as the FPGA development platform. Its Artix-7 fabric---240 DSP slices, approximately 4,860~Kb of block RAM, and 63,400 LUTs \cite{b11}---is large enough to implement a complete signal processing datapath but constrained enough that students must make deliberate resource allocation decisions. This is an important design choice: a platform with unlimited resources would hide the tradeoffs that make FPGA design interesting. Students learn to partition logic between the serial ingestion front-end, the packet parsing and CRC pipeline, and the feature extraction back-end within a fixed budget, which is the same kind of reasoning required in professional FPGA development. The board connects to the STM32 receiver through its PMOD headers, giving students a concrete physical interface to characterize and debug. Prior work has shown that inference pipelines of comparable complexity fit well within the resource envelope of similar devices \cite{b1,b3}, confirming that the Artix-7 is not a limiting factor for the scope of this curriculum.

At the analysis layer, a host machine accumulates the labeled feature vectors exported by the FPGA and serves as the ML development environment. Students select and train classifiers against this hardware-collected dataset, guided by prior comparative results on RF modulation classification \cite{b4} and LoRa-specific deep learning benchmarks \cite{b5} as reference points for evaluating their own models.

\subsection{Engagement Simulator}

To bridge the gap between curriculum introduction and physical hardware bring-up, a custom software engagement simulator was developed in Python as a standalone pre-hardware learning tool. The simulator models a multi-asset electronic warfare (EW) scenario---two cooperative target stations and an adversarial jamming threat---using a tick-based event engine that processes communication, jamming, and counter-jamming interactions at a configurable rate (500~ms per tick by default).

The simulator incorporates a physics-grounded RF model directly connected to the theoretical content students encounter in Module 1. Free-space path loss (FSPL) is computed per the standard Friis equation, and maximum communication range is derived from a link budget using configurable transmit power, antenna gain, and receiver sensitivity parameters. Target assets are modeled as frequency-hopping spread spectrum (FHSS) radios operating at 225~MHz (UHF military band) with eight available channels and a seed-synchronized hop sequence that both target stations share; the jamming threat uses a wideband receiver and DRFM-style noise jammer. This FHSS model connects directly to the spectral and synchronization concepts students work with during FPGA RTL design in Module 2, where clock domain management and frequency-selective filtering are primary design challenges.

Asset positions are specified in WGS-84 geodetic coordinates (latitude, longitude, altitude), with haversine-based range computation and geographic visualization on a 2-D map canvas. The default scenario places assets at coordinates representative of a realistic operational area, giving students immediate exposure to the link budget reasoning---how transmitted power, range, and receiver sensitivity interact---that governs the LoRa beacon configurations they will set in Module 1.

At each simulation tick, the engine resolves target-to-target communication attempts, threat jamming actions, and target counter-jamming responses in sequence. The outcome of each event is determined by the current channel alignment, effective range, and stochastic counter-jamming probability (default: 40\%). All events are written to a dual-log output: \texttt{engagement.log} records timestamped event descriptions, and \texttt{data\_stream.log} records the raw binary sequences exchanged---structured DATA, ACK, JAM, and counter-JAM patterns. These labeled logs serve as the first training dataset available to students in Module 4, allowing ML pipeline development to begin before any hardware RF data has been collected. Critically, the ground-truth labels (communication success, jamming state, counter-jamming outcome) are embedded directly in the log structure, giving students a clean supervised learning problem on which to validate preprocessing and model training code before the noisier, real-world dataset from the FPGA is available.

Per-tick performance metrics---communication success rate, jamming effectiveness, and counter-jamming rate---are accumulated over a rolling 60-tick history and displayed as live charts. These same three metrics form the primary benchmarks students measure on the physical system in Module 5, so the simulator establishes the measurement methodology and expected performance envelope before students have seen any hardware results.

\subsection{Curriculum Module Sequence}

\textit{Module 1: RF Hardware and Subsystem Design.} Students implement beacon transmission firmware on the STM32 platform, design the packet protocol and transmission scheduling algorithm, and validate the physical receiver-to-FPGA serial interface. The primary learning objective is understanding how firmware design choices---SPI clock rate, DMA configuration, interrupt latency---affect the timing and integrity of the data stream delivered to the FPGA. The Nucleo-L476RG's four independent serial interfaces and hardware-accelerated peripherals \cite{b10} provide enough configuration space to make these tradeoffs concrete and experimentally observable.

\textit{Module 2: FPGA RTL Design and Simulation.} Students implement the FPGA-side data ingestion pipeline in VHDL or Verilog: serial receiver, packet parser, CRC checker, and a feature extraction datapath. Simulation testbenches use recorded beacon data from Module 1 as stimulus. Students encounter FPGA-specific design constraints---timing closure, clock domain crossing, memory resource allocation---in the context of a real signal protocol rather than a synthetic exercise. Cloud-based FPGA toolchains have demonstrated that structured, hardware-grounded lab sequences can make this learning accessible without requiring specialized hardware beyond a web browser \cite{b9}; this module follows that model while adding the RF protocol context that a standalone digital design course omits.

\textit{Module 3: System Integration and Data Collection.} Students connect the hardware subsystems built in Modules 1 and 2 and characterize end-to-end data flow: from RF beacon transmission through reception and FPGA processing to feature output. Latency, throughput, and packet error rate are measured. Synchronization mismatches and protocol errors that were invisible in per-subsystem development surface here as integration failures that must be diagnosed and resolved. Students then execute a structured data collection campaign across multiple beacon configurations, producing a labeled RF dataset. This module deliberately surfaces the distribution shift problems that arise when training data is drawn from real hardware rather than simulation \cite{b6}.

\textit{Module 4: AI-Driven Classification Pipeline.} Using the hardware-collected dataset from Module 3, students preprocess FPGA-extracted features, train and evaluate classifiers, and analyze classification accuracy as a function of SNR and number of concurrent beacons. Streaming inference architectures \cite{b3} and FPGA-accelerated ML pipelines \cite{b2} serve as reference designs for students considering on-device deployment. Comparative architecture studies \cite{b4} and LoRa-specific classification results \cite{b5,b6,b8} provide empirical benchmarks against which students evaluate their own model performance.

\textit{Module 5: System Validation and Benchmarking.} Students stress-test the complete pipeline with multiple simultaneous beacons, measure total system response time from RF transmission to classification decision, and characterize the performance envelope of their implementation. The benchmarking exercise requires active engagement across all four domains: RF link budget analysis, FPGA timing and resource reports, serial interface throughput, and ML classification accuracy curves.

\subsection{Pedagogical Rationale}

The framework's central design principle is that the interfaces between disciplinary domains are the primary teaching objects, and this principle gives it concrete advantages over existing single-domain curricula. Prior work has demonstrated that hardware-grounded, progressive lab sequences are effective within a single discipline \cite{b9,b10}; this framework extends that validated model to four disciplines simultaneously. The key difference is that integration failures---a firmware DMA timing error that corrupts the FPGA data stream, or a beacon power change that degrades classifier accuracy---are not obstacles to be avoided but structured learning events. Each one forces the student to reason across domain boundaries in a way that no isolated course can replicate.

The framework is also designed to succeed where purely simulation-based approaches fall short. Prior work on LoRa signal classification consistently shows a gap between models trained on synthetic or benchmark data and those trained on real hardware captures \cite{b6,b8}. By staging the curriculum so that students first develop their ML pipeline on the engagement simulator's labeled synthetic data and then retrain on FPGA-collected hardware data in Module 4, the framework makes that distribution shift a deliberate teaching moment rather than a surprise. Students observe directly how channel impairments, timing offsets, and hardware non-idealities affect model performance---a lesson that cannot be delivered on a curated academic dataset.

Finally, the framework is designed for reproducibility and transferability. All components---the Nexys A7-100T, the STM32 Nucleo-L476RG, the RFM95W, and the engagement simulator---are commercially available or open-source. Prior work has validated that the core technical pipelines are feasible within these hardware resource budgets \cite{b1,b3,b7}, meaning an instructor at another institution can deploy the same curriculum without custom hardware or proprietary infrastructure. The framework's contribution is not a novel algorithm or hardware design but a structured, teachable model of how these four engineering disciplines connect at the system level---with each connection made concrete through implementation on real hardware.

\begin{thebibliography}{00}

\bibitem{b1}
H. Komkov, L. Pocher, A. Restelli, B. Hunt, and D. Lathrop, ``RF signal classification using Boolean reservoir computing on an FPGA,'' in \textit{Proc. 2021 Int. Joint Conf. Neural Networks (IJCNN)}, Jul. 2021, doi: 10.1109/IJCNN52387.2021.9533342.

\bibitem{b2}
J. F. Lin, T. Morehouse, C. Montes, E. Caushi, A. Dudko, E. Savage, and R. Zhou, ``A study of angle of arrival estimation of a RF signal with FPGA acceleration,'' in \textit{Proc. 2024 Int. Wireless Commun. Mobile Comput. Conf. (IWCMC)}, 2024, pp. 1625--1628, doi: 10.1109/IWCMC61514.2024.10592551.

\bibitem{b3}
A. Maclellan, L. H. Crockett, and R. W. Stewart, ``Streaming convolutional neural network FPGA architecture for RFSoC data converters,'' in \textit{Proc. 2023 21st IEEE Interreg. NEWCAS Conf. (NEWCAS)}, Edinburgh, UK, Jun. 2023, doi: 10.1109/NEWCAS57931.2023.10198198.

\bibitem{b4}
K. A. Alnajjar, S. Ghunaim, and S. Ansari, ``Automatic modulation classification in deep learning,'' in \textit{Proc. 2022 5th Int. Conf. Commun., Signal Process., Appl. (ICCSPA)}, Cairo, Egypt, Dec. 2022, doi: 10.1109/ICCSPA55860.2022.10019122.

\bibitem{b5}
P. M. Mutescu, A. Lavric, A. I. Petrariu, and V. Popa, ``Deep learning enhanced spectrum sensing for LoRa spreading factor detection,'' in \textit{Proc. 2023 13th Int. Symp. Adv. Topics Elect. Eng. (ATEE)}, Bucharest, Romania, Mar. 2023, doi: 10.1109/ATEE58038.2023.10108224.

\bibitem{b6}
Y. E. Sagduyu and T. Erpek, ``Adversarial attacks on LoRa device identification and rogue signal detection with deep learning,'' in \textit{Proc. MILCOM 2023 --- 2023 IEEE Military Commun. Conf.}, Boston, MA, Oct.--Nov. 2023, pp. 385--390, doi: 10.1109/MILCOM58377.2023.10356224.

\bibitem{b7}
T.-P. Dang, T.-K. Tran, T.-T. Bui, and H.-T. Huynh, ``LoRa gateway based on SoC FPGA platforms,'' in \textit{Proc. 2021 Int. Symp. Elect. Electron. Eng. (ISEE)}, 2021, pp. 48--52, doi: 10.1109/ISEE51682.2021.9418711.

\bibitem{b8}
K. Dakic, B. Al Homssi, A. Al-Hourani, and M. Lech, ``LoRa signal demodulation using deep learning, a time-domain approach,'' in \textit{Proc. 2021 IEEE 93rd Veh. Technol. Conf. (VTC2021-Spring)}, Apr. 2021, doi: 10.1109/VTC2021-Spring51267.2021.9448711.

\bibitem{b9}
W. Smith, Z. Driskill, J. Goeders, and M. Wirthlin, ``Digital design education using an open-source, cloud-based FPGA toolchain,'' in \textit{Proc. 2024 Intermountain Eng., Technol. Comput. Conf. (IETC)}, May 2024, doi: 10.1109/IETC61393.2024.10564285.

\bibitem{b10}
N. O. Strelkov, V. V. Krutskikh, and E. V. Shalimova, ``Programming STM32 Nucleo platform for IoT education using STM32duino and Mbed OS,'' in \textit{Proc. 2022 VI Int. Conf. Inf. Technol. Eng. Educ. (Inforino)}, Apr. 2022, doi: 10.1109/Inforino53888.2022.9782920.

\bibitem{b11}
Digilent Inc., ``Nexys A7 FPGA Board Reference Manual,'' Digilent Inc., Pullman, WA, 2023. [Online]. Available: https://digilent.com/reference/programmable-logic/nexys-a7/reference-manual

\bibitem{b12}
Adafruit Industries, ``Adafruit RFM69HCW and RFM9X LoRa Packet Radio Breakouts,'' Adafruit Industries, New York, NY, 2023. [Online]. Available: https://learn.adafruit.com/adafruit-rfm69hcw-and-rfm96-rfm95-rfm98-lora-packet-padio-breakouts/downloads

\bibitem{b13}
STMicroelectronics, ``STM32 Nucleo-64 Boards (MB1136) User Manual,'' UM1724, STMicroelectronics, Geneva, Switzerland, 2023. [Online]. Available: https://www.st.com/en/evaluation-tools/nucleo-l476rg.html

\end{thebibliography}

\end{document}
