\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-E}
\newcommand{\labtitleshort}{Failure Mode Catalog}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-E}{Failure Mode Catalog and Cross-Domain Dependency Map}

\section{Objective}
Assemble the structured catalog of every failure or degradation mode observed across Modules~1--5, trace each entry back to the originating design decision, and produce the cross-domain dependency diagram that visualises how a decision in one module propagated into a failure in another. This handout is the project's most explicit statement of the curriculum's core pedagogical claim: engineering systems are defined by their interfaces, not their components.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handouts~5-A through~5-D complete (all stress/benchmark results in hand, including the S5 PAUSE behavioural-response intervals from LH5-C).
	\item Access to all prior handouts' deliverables so every anomaly tagged during Modules~1--4 can be revisited.
\end{itemize}

\section{Procedure}

\subsection{Step 0 --- The running log (should have started in Module~1)}

\begin{tcolorbox}[warning={Start this file at the beginning of the project, not here.}]
	The failure mode catalog is most valuable when entries are written \emph{at the time the anomaly is observed}, not reconstructed weeks later from memory. If you have not been keeping a running log, start now and back-fill what you can recall. Future iterations of this curriculum should begin maintaining \texttt{failure\_modes.csv} from LH1-D onward.
\end{tcolorbox}

Maintain a file \texttt{failure\_modes.csv} at the root of the project repository throughout all five modules. Its schema:

\begin{lstlisting}[caption={\texttt{failure\_modes.csv} --- append one row per observed anomaly.}]
id,failure_mode,observed_in,root_module,root_cause,depth,status
F1,FIFO overflow under burst arrivals,LH5-A S1c/e,M2,FIFO depth 64 insufficient for two-Target ALOHA + PAUSE peaks,1,documented
F2,CRC reject on valid frames,LH3-C,M1/M2,CRC polynomial mismatch between firmware and RTL,1,fixed
F3,Classification collapse at low SNR,LH5-B S3d/e,M3,Training data did not cover <0 dB SNR regime,2,documented
\end{lstlisting}

Column definitions:
\begin{itemize}[leftmargin=2em]
	\item \textbf{id}: F1, F2, \ldots (assigned sequentially as anomalies are observed; do not renumber).
	\item \textbf{failure\_mode}: one-line description using the system's vocabulary.
	\item \textbf{observed\_in}: the handout + step where it was first seen (e.g.\ \texttt{LH3-C step 4}).
	\item \textbf{root\_module}: the \emph{earliest} module whose design decision is responsible.
	\item \textbf{root\_cause}: the specific causal chain from root decision to observed symptom. Name the specific parameter (e.g.\ ``FIFO depth 64''), not a broad category.
	\item \textbf{depth}: 0 = self-contained, 1 = one module boundary crossed, 2+ = two or more.
	\item \textbf{status}: \texttt{fixed} / \texttt{documented} / \texttt{mitigated} / \texttt{open}.
\end{itemize}

To validate the CSV at any point during the project:

\begin{lstlisting}[language=Python, caption={\texttt{validate\_failure\_log.py} --- run from project root; prints PASS or lists missing fields.}]
import csv, sys
from pathlib import Path

REQUIRED = {"id", "failure_mode", "observed_in", "root_module",
            "root_cause", "depth", "status"}
VALID_DEPTH  = {"0", "1", "2+"}
VALID_STATUS = {"fixed", "documented", "mitigated", "open"}

rows = list(csv.DictReader(Path("failure_modes.csv").open()))
errors = []

for r in rows:
    missing = REQUIRED - set(r.keys())
    if missing:
        errors.append(f"{r.get('id','?')}: missing fields {missing}")
    if r.get("depth") not in VALID_DEPTH:
        errors.append(f"{r['id']}: depth must be 0/1/2+, got '{r['depth']}'")
    if r.get("status") not in VALID_STATUS:
        errors.append(f"{r['id']}: status must be fixed/documented/mitigated/open")

depth2 = sum(1 for r in rows if r.get("depth") == "2+")
if len(rows) < 8:
    errors.append(f"Only {len(rows)} entries; need >= 8")
if depth2 < 2:
    errors.append(f"Only {depth2} depth-2+ entries; need >= 2")

if errors:
    print("FAIL")
    for e in errors:
        print(" ", e)
    sys.exit(1)
print(f"PASS  {len(rows)} entries, {depth2} depth-2+")
\end{lstlisting}

\subsection{Step 1 --- Populate the catalog}
For every failure, integration issue, or accuracy anomaly observed during the project (yes, including ones that were fixed), create one catalog entry. Aim for $\geq$ 8 entries. The framework's pedagogical depth is directly proportional to the richness of this list. The starter table below mirrors the F1--F8 catalog from the Module~5 lesson plan; replace each row with your own evidence and refine the root-cause language.

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c l c c X}
		\toprule
		\textbf{ID} & \textbf{Failure Mode} & \textbf{Observed In} & \textbf{Root Module} & \textbf{Root Cause} \\
		\midrule
		F1 & \FIFO{} overflow under burst arrivals             & LH5-A (S1c, S1e) & M2   & \FIFO{} depth 64 insufficient for two-Target ALOHA + intermittent PAUSE peaks \\
		F2 & \mbox{CRC} reject on valid frames                  & LH3-C            & M1/M2 & \mbox{CRC} polynomial mismatch between firmware and \mbox{RTL} \\
		F3 & Classification collapse at low \mbox{SNR}          & LH5-B (S3d/e)    & M3/M4 & Training data did not cover $<$0\,dB \mbox{SNR} regime \\
		F4 & Transient SF/mod misclass at SF transition         & LH5-B (S2)       & M2   & \texttt{inter\_arrival\_ms} computed against previous-SF DATA frame from same Target~ID \\
		F5 & Stale RSSI table $\Rightarrow$ wrong PAUSE target  & LH5-C (S5)       & M1   & Threat RSSI table not aged; old strongest emitter selected after the link rebalances \\
		F6 & Both Targets simultaneously jammed                 & LH5-A (S1d)      & M1   & Held-button PAUSE without RSSI tie-break selects the same Target every cadence; the other never receives a PAUSE row but is collateral-jammed by the on-air burst \\
		F7 & Operating-condition shift (clean$\to$contended)    & LH4-F + LH5-D B5 & M3   & Training set under-samples PAUSE-active regime; classifier accuracy drops once C5 introduces PAUSE rows and contention \\
		F8 & Host stdout buffering inflates $T_{\text{infer}}$  & LH5-C            & M4   & \texttt{print()} buffering measured inside the timed region \\
		\bottomrule
	\end{tabularx}
	\caption{Starter catalog --- replace each row with your own observed failure data. Each entry must cite the specific handout in which it was observed.}
\end{table}

\begin{tcolorbox}[infobox={Catalog Entry Structure}]
	Each entry must contain five fields:
	\begin{enumerate}[leftmargin=1.2em]
		\item \textbf{ID:} F1, F2, ... (unique).
		\item \textbf{Failure mode:} a one-line description using the system's vocabulary (not a narrative).
		\item \textbf{Observed in:} the handout + step where it was first seen.
		\item \textbf{Root module:} the \emph{earliest} module whose design decision is responsible. This is the cross-domain insight; it is almost never the same as ``observed in''.
		\item \textbf{Root cause:} a one- or two-sentence explanation of the causal chain from the root design decision to the observed symptom. Include the \emph{specific} design parameter (e.g., ``\FIFO{} depth 64''), not just the broad category (``memory size'').
	\end{enumerate}
\end{tcolorbox}

\subsection{Step 2 --- Classify each entry by chain depth}
Tag each catalog entry with its \emph{chain depth}: the number of module boundaries the causal chain crosses.
\begin{itemize}[leftmargin=2em]
	\item \textbf{Depth 0:} Root module = Observed module. A failure internal to one subsystem.
	\item \textbf{Depth 1:} Root module is the immediate upstream neighbour (common: M1 causes something in M2).
	\item \textbf{Depth 2+:} Cross-module-skipping chain (e.g., M1 firmware decision surfaces as an M4 accuracy problem). These are the pedagogically richest entries.
\end{itemize}

Count the distribution. A healthy catalog has at least a couple of depth-2+ entries --- if every entry is depth 0, the integration surface was too thin, suggesting the curriculum could be more interconnected. F5 (M1$\to$M5) and F7 (M3$\to$M5) are good candidate depth-2+ entries from the starter set.

\subsection{Step 3 --- Cross-domain dependency diagram}
Draw a directed graph using TikZ (or any diagramming tool). Nodes are design decisions; edges point from ``decision'' to ``observed failure''. Group nodes by module.

\begin{lstlisting}[language={[LaTeX]TeX}, caption={Starter TikZ skeleton for \texttt{cross\_domain\_map.tex} --- extend with your own nodes.}]
\begin{tikzpicture}[
  node distance=1.2cm and 1.6cm,
  decision/.style={rectangle, draw=stevens, fill=lightgray,
                   rounded corners, align=center, font=\small,
                   minimum width=3cm, minimum height=1cm},
  failure/.style={rectangle, draw=accentblue, fill=blue!5,
                  align=center, font=\small,
                  minimum width=3cm, minimum height=1cm},
  arrow/.style={-{Stealth[length=3mm]}, thick}
]
  % Module 1 design decisions
  \node[decision] (baud)  {UART 115200 baud\\(M1)};
  \node[decision, right=of baud] (pkt)   {32-B packet,\\packed Mod-Code\\(M1)};
  \node[decision, right=of pkt]  (rssi)  {Threat RSSI table\\age policy (M1)};

  % Module 2 design decisions
  \node[decision, below=of baud] (fifo)  {FIFO depth 64\\(M2)};
  \node[decision, below=of pkt]  (feat)  {feature vector w/\\pkt\_type @ off.~3\\(M2)};

  % Module 3 design decision
  \node[decision, below=of rssi] (regime) {C1--C4 only in\\training set (M3)};

  % Observed failures
  \node[failure, below=of fifo]   (F1) {F1: FIFO overflow\\(LH5-A S1c/e)};
  \node[failure, below=of feat]   (F4) {F4: Transient miscl.\\at SF transition};
  \node[failure, below=of regime] (F7) {F7: op-shift drop\\(LH4-F, LH5-D B5)};
  \node[failure, right=of F7]     (F5) {F5: stale RSSI\\$\Rightarrow$ wrong PAUSE\\target (LH5-C S5)};

  \draw[arrow] (baud)   -- (fifo);
  \draw[arrow] (fifo)   -- (F1);
  \draw[arrow] (feat)   -- (F4);
  \draw[arrow] (pkt)    -- (F4);
  \draw[arrow] (regime) -- (F7);
  \draw[arrow] (rssi)   -- (F5);
  % ... extend ...
\end{tikzpicture}
\end{lstlisting}

Goal: the reader should be able to trace any failure node back to its root module by following the arrows backward. This is the visual counterpart to the catalog table.

\subsection{Step 4 --- Depth-2 walkthrough}
Pick one depth-2 or deeper entry from your catalog and write a one-page walkthrough of the full causal chain. Example structure for F7 (``Operating-condition shift, clean$\to$contended''):

\begin{enumerate}[leftmargin=2em]
	\item \textbf{Design decision (M3):} The Module~3 collection campaigns C1--C4 were all run in the ``clean'' regime --- no PAUSE asserted, no contention beyond the natural Target~A / Target~B ping-pong. Chosen for reproducibility and labelling simplicity at 906.5\,MHz.
	\item \textbf{Propagation (M3 dataset):} The training set therefore has near-zero PAUSE-row examples and near-zero half-jammed feature vectors. The marginal distributions of \texttt{rssi\_delta}, \texttt{inter\_arrival\_ms}, and the packed Modulation Code byte are all skewed toward the clean-regime mode.
	\item \textbf{Propagation (M4):} The four-task classifier learned class-conditional distributions on this clean training set. The Module~4 LH4-F operating-condition shift experiment quantified $\Delta_{\text{op-shift}}$ on a held-out C5 slice and found the largest drops on the modulation-kind and packet-type tasks.
	\item \textbf{Observed failure (M5):} In LH5-D's B5 operating point (Target~A + Target~B + held-button PAUSE), the same drop reappears live. PAUSE rows are sometimes misclassified as DATA, and modulation-kind accuracy on the half-suppressed Target degrades by $>$5\,pp from the B4 baseline.
	\item \textbf{Fix candidates (M3 retrofit):} Add a dedicated C5-style collection campaign with intermittent PAUSE asserted, or apply targeted oversampling/fine-tuning on the small contended slice (the LH4-F fine-tuning recipe).
\end{enumerate}

The walkthrough is not a generic post-mortem --- it is specifically an articulation of \emph{how the earliest module's seemingly-reasonable choice constrained the latest module's performance}. That articulation is the point of the whole framework.

\subsection{Step 5 --- Known-limitations summary}
From the catalog, distil a ``Known Limitations'' section (the final report will cite it verbatim):

\begin{itemize}[leftmargin=2em]
	\item Operating envelope boundaries identified empirically (max active stations, min \mbox{SNR}, max distance, PAUSE response/recovery latency).
	\item Failure modes not fully mitigated (cite F-IDs).
	\item Cross-module dependencies that would require a redesign to eliminate (e.g., F7 needs a Module~3 dataset expansion, not a Module~4 fix).
\end{itemize}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] \texttt{validate\_failure\_log.py} prints \texttt{PASS} (8+ entries, 2+ depth-2+, all fields valid).
	\item[$\square$] Each entry has a depth tag (0, 1, 2+).
	\item[$\square$] At least two depth-2+ entries present (if fewer, add cross-module re-analysis of existing observations).
	\item[$\square$] Cross-domain dependency diagram generated and legible.
	\item[$\square$] One-page walkthrough of a depth-2 chain written.
	\item[$\square$] Known Limitations summary written.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-E Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item \texttt{failure\_modes.csv} with $\geq$8 entries, all seven columns populated, \texttt{validate\_failure\_log.py} printing \texttt{PASS}.
		\item Failure mode catalog table (can be generated from the CSV or filled directly in LaTeX).
		\item Cross-domain dependency diagram (PDF or PNG).
		\item One-page depth-2 walkthrough tracing a single failure through the full causal chain.
		\item Known Limitations summary (bulleted, fits in half a page).
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
