\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-C}
\newcommand{\labtitleshort}{End-to-End Latency + S5 PAUSE Response}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-C}{End-to-End Latency Instrumentation, Per-Stage Decomposition, and S5 PAUSE Behavioural Response}

\section{Objective}
Instrument the full pipeline --- Target~A \mbox{TX} start $\to$ Threat interceptor classification output --- with a GPIO-toggle latency measurement and produce a per-stage decomposition of the total response time. This handout extends the LH3-D \mbox{GPIO}-toggle method from a single stage (\mbox{RF} + firmware) to all six stages, giving the Module~5 benchmarking table in LH5-D a physical measurement per stage rather than a theoretical budget. The host-side ``classification output'' is the three-task (\texttt{sf}, \texttt{beacon}, \texttt{pkt}) prediction emitted by \texttt{live\_inference.py}; the dropped \texttt{mod} task does not contribute to $T_{\text{infer}}$. This handout also runs the \textbf{S5 PAUSE behavioural-response} stress test, which measures the closed-loop latency from the held-button assertion at the Threat to the moment a Target prints \texttt{I'm Jammed} on USART2, and the recovery latency once the button is released.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item LH3-D complete (\mbox{GPIO}-toggle method validated for Target$\to$\FPGA{} latency).
	\item Two-channel oscilloscope or logic analyzer available (four channels strongly preferred for S5).
	\item USB-\mbox{GPIO} adapter (e.g., Arduino, FT232H, or Raspberry Pi Pico) for the host-side toggle. If unavailable, an alternative using \texttt{frame\_valid} as the endpoint is documented in Step~5.
	\item Target~A, Target~B, and the Threat interceptor all operational on 906.5\,MHz, with the PAUSE button wired to the Threat's user push-button (PC13) per LH1-G.
\end{itemize}

\section{Latency Model}
\[
	T_{\text{total}} = T_{\text{RF}} + T_{\text{RX\_FW}} + T_{\text{UART\_in}} + T_{\text{FPGA}} + T_{\text{UART\_out}} + T_{\text{infer}}
\]

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l c l X}
		\toprule
		\textbf{Stage} & \textbf{Symbol} & \textbf{Method} & \textbf{Description} \\
		\midrule
		\mbox{RF} on-air          & $T_{\text{RF}}$       & Calculated from SF/BW & LoRa packet duration on the air \\
		Threat firmware            & $T_{\text{RX\_FW}}$    & \mbox{GPIO} toggle     & \mbox{ISR} + 43-byte frame construction in the Threat STM32 \\
		\mbox{UART} to \FPGA{}     & $T_{\text{UART\_in}}$  & Calculated            & $43 \times 10 / 115{,}200 = 3.73$\,ms \\
		\FPGA{} pipeline           & $T_{\text{FPGA}}$      & \mbox{ILA}            & UART-RX-complete to \texttt{frame\_valid} \\
		\mbox{UART} to host        & $T_{\text{UART\_out}}$ & Calculated            & $35 \times 10 / 115{,}200 = 3.04$\,ms \\
		Host inference (3 tasks)   & $T_{\text{infer}}$     & \texttt{perf\_counter()} & Parse + scale + three sequential predicts (\texttt{sf}, \texttt{beacon}, \texttt{pkt}) \\
		\bottomrule
	\end{tabularx}
\end{table}

\section{Procedure}

\subsection{Step 1 --- Target-side GPIO toggle}
From LH3-D, Target~A already toggles PA8 (or the board \mbox{LED} on PA5) just before \texttt{sx1276\_send()}. Confirm this is still present in the firmware. If not, re-add the single-line \texttt{HAL\_GPIO\_WritePin} immediately before the \mbox{SPI} transmit that initiates the \mbox{TX}. Target~B should have the same toggle on the same pin so either DATA or ACK frames can be used as the trigger.

\subsection{Step 2 --- Host-side GPIO toggle}
Wire a USB-\mbox{GPIO} adapter to the host. In \texttt{live\_inference.py}, add a one-line pin-toggle call on every classification print:

\begin{lstlisting}[language=Python, caption={Host-side toggle snippet --- add just before the \texttt{print()} in the live loop. Note the three-task field set (the deployed pipeline does not include a \texttt{mod} task).}]
# gpio is a handle to a USB-GPIO adapter (Arduino-Firmata, FT232H, pyftdi...)
gpio.toggle(HOST_LATENCY_PIN)   # 1 us resolution via FT232H
print(f"[{ts}] SF={preds['sf']['label']} "
      f"BEACON={preds['beacon']['label']} "
      f"PKT={preds['pkt']['label']}")
\end{lstlisting}

The toggle should fire on the \emph{first line} of the output path, not inside the \texttt{print()} call, so that Python stdout buffering does not inflate $T_{\text{infer}}$.

\subsection{Step 3 --- Scope capture}
Connect:
\begin{itemize}[leftmargin=2em]
	\item Ch.~1: Target~A \mbox{GPIO} toggle (rising edge marks DATA \mbox{TX} start).
	\item Ch.~2: host \mbox{GPIO} toggle (rising edge marks three-task classification output).
	\item Ch.~3 (optional, if a 4-channel scope is available): \FPGA{} \texttt{frame\_valid} routed to \mbox{PMOD} JA[2] (as in LH3-D).
\end{itemize}

Trigger on Ch.~1 rising. Capture at least 50 consecutive events per operating point.

\subsection{Step 4 --- Per-stage decomposition}
Combine measurements and calculations per operating point:

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l X l}
		\toprule
		\textbf{Measurement} & \textbf{How to derive} & \textbf{Required uncertainty} \\
		\midrule
		$T_{\text{total}}$        & Ch.~2 rising edge $-$ Ch.~1 rising edge (scope cursor) & $\pm$(scope cursor resolution + trigger jitter) \\
		$T_{\text{RF}}$           & Theoretical LoRa on-air time for the configured SF/BW/payload & $\pm$(SX1276 symbol-timing tolerance, typ.\ 1\,\%) \\
		$T_{\text{FPGA}}$         & \mbox{ILA} trace: UART-RX-complete $\to$ \texttt{frame\_valid} (LH3-C method) & $\pm$(1 sysclk cycle, i.e.\ 10\,ns at 100\,MHz) \\
		$T_{\text{UART\_in}}$ / $T_{\text{UART\_out}}$ & Fixed: 3.73\,ms / 3.04\,ms at 115200 baud & $\pm$(baud-divisor quantization, typ.\ 0.5\,\%) \\
		$T_{\text{infer}}$        & \texttt{time.perf\_counter()} around the three-task \texttt{InferenceEngine.predict()} call (LH4-H microbench --- the deployed mix is CNN(sf) + SVM(beacon) + SVM(pkt)) & $\pm$(reported std over $\geq$100 calls) \\
		$T_{\text{RX\_FW}}$       & $T_{\text{total}}$ $-$ (all other stages above); residual & \textbf{Required:} quadrature-sum of all other stages' uncertainties \\
		\bottomrule
	\end{tabularx}
	\caption{Per-stage decomposition with required uncertainty bounds. Three stages are directly measured, three are calculated or inferred by subtraction. \textbf{Each row must carry an explicit uncertainty} --- instrument resolution for measured stages, tolerance or quantization for calculated stages, and the propagated residual for $T_{\text{RX\_FW}}$. Without uncertainty bounds the decomposition is not defensible as a spec claim in LH5-D.}
\end{table}

\begin{tcolorbox}[warning={Uncertainty Propagation for the Residual Stage}]
The residual term $T_{\text{RX\_FW}} = T_{\text{total}} - \sum_{i \neq \text{RX\_FW}} T_i$ inherits the uncertainty of every stage subtracted from it. Report it as
\[
\sigma_{\text{RX\_FW}} = \sqrt{\sigma_{\text{total}}^2 + \sum_{i \neq \text{RX\_FW}} \sigma_i^2}.
\]
If $\sigma_{\text{RX\_FW}}$ is comparable to the residual value itself, state clearly that the Threat-firmware stage is ``not resolvable by subtraction at this instrumentation budget'' and do not present a point estimate without the bound.
\end{tcolorbox}

\subsection{Step 5 --- Alternative if no host USB-GPIO is available}
If a USB-\mbox{GPIO} adapter cannot be obtained, use the \FPGA{}'s \texttt{frame\_valid} (routed to \mbox{PMOD} JA[2] per LH3-D) as the measurement endpoint:

\[
	T_{\text{total}}^{(\text{partial})} = T_{\text{RF}} + T_{\text{RX\_FW}} + T_{\text{UART\_in}} + T_{\text{FPGA}}
\]

Then add the host-side contributions ($T_{\text{UART\_out}} + T_{\text{infer}}$) from the LH4-H microbenchmark. Document this substitution explicitly in the report --- the scope-measured envelope and the host-measured envelope are then added \emph{analytically}, not through a single end-to-end trigger. The p99 total may be slightly under-reported because this method captures the typical host latency rather than the exact tail of any given transaction.

\subsection{Step 6 --- Verification against LH3-D}
At SF7, the scope-measured $T_{\text{total}} - T_{\text{infer}} - T_{\text{UART\_out}}$ should match the LH3-D Step~4 measurement (mean near 55\,ms, std $<$5\,ms). If it does not, one of the instrumentation paths has drifted --- re-check pin assignments, trigger level, and the Target firmware \mbox{GPIO} toggle polarity.

\begin{tcolorbox}[infobox={Why the RF Stage Dominates}]
	The \mbox{RF} on-air time is $\sim$50\,ms at SF7 and $\sim$1.5\,s at SF12, while every downstream stage is $<$10\,ms combined. The dominant stage is \emph{always} $T_{\text{RF}}$ at any LoRa SF. This has an important architectural consequence: moving the three-task inference step onto the \FPGA{} would eliminate $T_{\text{UART\_out}} + T_{\text{infer}}$ ($\sim$5--6\,ms with the deployed CNN+SVM+SVM mix), which is a $<$12\% reduction in $T_{\text{total}}$ at SF7 and $<$1\% at SF12. The investment is only worthwhile if the host is also eliminated (standalone embedded deployment) --- a point LH5-D asks you to articulate.
\end{tcolorbox}

\subsection{Step 7 --- S5: PAUSE Behavioural Response Latency}
With the latency rig still in place, run the closed-loop S5 measurement. Configure Target~A and Target~B in fixed 1-second DATA/ACK ping-pong at SF7, +14\,dBm. The Threat is powered, listening, and the operator is ready to assert PAUSE.

Re-route the scope channels for this step:
\begin{itemize}[leftmargin=2em]
	\item Ch.~1: Threat user-button input (PC13). Falling edge marks the operator pressing PAUSE.
	\item Ch.~2: Threat \mbox{TX} indicator (toggle the same PA8/PA5 line just before \texttt{sx1276\_send\_pause()} in the LH1-G PAUSE scheduler). Rising edge marks the first PAUSE burst leaving the Threat.
	\item Ch.~3: Target~A or Target~B USART2 \mbox{TX} pin (the line that emits \texttt{"I'm Jammed\textbackslash r\textbackslash n"} per LH1-F). The first start bit (falling edge) marks the Target's behavioural acknowledgement of PAUSE.
	\item Ch.~4 (optional): host \mbox{GPIO} toggle from Step~2, so the three-task classifier's first \texttt{pkt=PAUSE} print is also captured.
\end{itemize}

Define and capture the four S5 latency intervals:

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l l l X}
		\toprule
		\textbf{Interval} & \textbf{Start edge} & \textbf{End edge} & \textbf{Meaning} \\
		\midrule
		$T_{\text{btn}\to\text{TX}}$    & Ch.~1 falling (button press) & Ch.~2 rising (first PAUSE burst) & Threat reaction time: RSSI-table lookup + scheduler wake \\
		$T_{\text{TX}\to\text{ACK}}$    & Ch.~2 rising (PAUSE burst)   & Ch.~3 falling (Target USART2 first start bit) & On-air + Target firmware response \\
		$T_{\text{btn}\to\text{cls}}$   & Ch.~1 falling                & Ch.~4 rising (\texttt{pkt=PAUSE} print) & End-to-end classifier acknowledgement \\
		$T_{\text{recover}}$            & Ch.~1 rising (button release)& Next Target~A DATA frame back on air (Ch.~3 high) & Hold-off expiry + return to ping-pong \\
		\bottomrule
	\end{tabularx}
\end{table}

Capture a minimum of 30 button-press / button-release cycles. Vary the operator dwell time across cycles (250\,ms, 1\,s, 5\,s) so the recovery distribution covers both ``hold-off dominated'' and ``pure RF re-sync'' regimes.

\subsection{Step 8 --- S5 analysis}
Compute mean, std, and p99 for each of the four intervals. Pre-declared S5 thresholds (mirrored from the spec table in LH5-D):
\begin{itemize}[leftmargin=2em]
	\item $T_{\text{btn}\to\text{TX}}$ p99 $\leq$ 100\,ms (one PAUSE-cadence period).
	\item $T_{\text{btn}\to\text{ACK}} = T_{\text{btn}\to\text{TX}} + T_{\text{TX}\to\text{ACK}}$ p99 $\leq$ 250\,ms (one PAUSE-cadence + one SF7 round trip).
	\item $T_{\text{btn}\to\text{cls}}$ p99 $\leq$ 350\,ms (above plus a three-task inference cycle).
	\item $T_{\text{recover}}$ p99 $\leq$ 500\,ms (the LH1-F default hold-off).
\end{itemize}

Report any cycle that misses these thresholds and root-cause it: PAUSE-cadence quantisation, RSSI-table staleness (failure mode F5 from LH5-E), or contention with concurrent ACK transmissions.

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] Target-side \mbox{GPIO} toggle firing on every \mbox{TX} (DATA and ACK).
	\item[$\square$] Host-side \mbox{GPIO} toggle firing on every classification (or alternative method documented).
	\item[$\square$] At least 50 $T_{\text{total}}$ measurements captured at SF7 baseline.
	\item[$\square$] Per-stage decomposition computed and checked for sum consistency ($\sum$ stages $=$ $T_{\text{total}}$).
	\item[$\square$] Cross-check against LH3-D confirms no drift in the Target$\to$\FPGA{} envelope.
	\item[$\square$] S5: at least 30 PAUSE press/release cycles captured across three dwell-time settings.
	\item[$\square$] S5: all four intervals ($T_{\text{btn}\to\text{TX}}$, $T_{\text{TX}\to\text{ACK}}$, $T_{\text{btn}\to\text{cls}}$, $T_{\text{recover}}$) summarised with p99.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-C Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item Oscilloscope captures: Ch.~1 (Target~A \mbox{TX}), Ch.~2 (host output), and optionally Ch.~3 (\texttt{frame\_valid}).
		\item Per-stage decomposition table at SF7 baseline (six rows).
		\item Sum-consistency check: $T_{\text{RF}} + T_{\text{RX\_FW}} + T_{\text{UART\_in}} + T_{\text{FPGA}} + T_{\text{UART\_out}} + T_{\text{infer}} \approx T_{\text{total}}$ within measurement uncertainty.
		\item Cross-check paragraph showing the LH5-C Target$\to$\FPGA{} envelope matches the LH3-D measurement within tolerance.
		\item S5 PAUSE behavioural response: scope captures, four-interval table with mean / std / p99, threshold-conformance verdict, and a one-paragraph root cause for any out-of-spec cycles.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
