\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-H}
\newcommand{\labtitleshort}{Presentation and Live Demo}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-H}{Final Presentation, Live Demonstration, and Capstone Submission}

\section{Objective}
Prepare and deliver the 20-minute capstone presentation (plus 10 minutes of Q\&A) that communicates the system architecture, experimental results, and lessons learned; include a live demonstration of real-time four-task classification and the button-gated PAUSE behavioural response; and submit the complete curriculum package. This is the final handout of the capstone --- all prior deliverables feed this one.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handout~5-G complete (final report written and reviewed).
	\item Lab Handout~5-F complete (curriculum package assembled and verified).
	\item All hardware operational; the bidirectional Target~A / Target~B ping-pong runs end-to-end without manual intervention, the Threat interceptor forwards 43-byte frames to the FPGA reliably, and the PAUSE button on the Threat triggers ``I'm Jammed'' on the targeted Target's USART2 within the LH5-C S5 thresholds.
\end{itemize}

\section{Presentation Requirements}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l X}
		\toprule
		\textbf{Parameter} & \textbf{Requirement} \\
		\midrule
		Duration   & 20 minutes presentation $+$ 10 minutes \mbox{Q\&A} \\
		Audience   & Project advisor, program committee, peers \\
		Format     & Slide deck (PowerPoint, Beamer, Google Slides) \\
		Live demo  & 3--4 minute live real-time classification + PAUSE button-hold demo, with a backup video \\
		\bottomrule
	\end{tabularx}
\end{table}

\section{Procedure}

\subsection{Step 1 --- Slide outline}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c l c X}
		\toprule
		\textbf{\#} & \textbf{Slide} & \textbf{Count} & \textbf{Contents} \\
		\midrule
		1 & Title                 & 1 & Project title, student, advisor, institution, date \\
		2 & Motivation \& goals   & 2 & Why this framework? What cross-domain skills does it teach? \\
		3 & System architecture   & 2 & Full system diagram (Target~A, Target~B, Threat); hardware photo; data-flow overview \\
		4 & Module highlights     & 5 & 1 slide per module: key decision, key result, link to next module \\
		5 & Live demo             & 1 + live & Real-time four-task classification; PAUSE button-hold behavioural response \\
		6 & Key results           & 3 & System spec table; per-task classification accuracy; stress-test findings; PAUSE response/recovery latency \\
		7 & Lessons learned       & 2 & Cross-domain dependencies; operating-condition shift; redesign choices \\
		8 & Conclusion \& future  & 1 & Summary, next steps \\
		9 & \mbox{Q\&A}           & 1 & Contact info, acknowledgments \\
		\bottomrule
	\end{tabularx}
	\caption{Target $\sim$18--20 slides. Rehearse to pace $\sim$1 minute per slide.}
\end{table}

\subsection{Step 2 --- Prepare each module slide}
The five module-highlight slides are the hardest to get right. For each one:
\begin{itemize}[leftmargin=2em]
	\item Pick the \emph{single} most compelling figure or table from that module.
	\item State in one phrase (on the slide) the key design decision: e.g., ``32-B packet, packed Modulation Code byte'', ``feature vector with packet-type at offset 3'', ``5 sweep campaigns, regime-stratified split'', ``four-task classifier suite at 100\,MHz'', ``button-gated PAUSE for behavioural response''.
	\item State the key result in a sentence or a number.
	\item \emph{In your script}: end with a sentence that transitions to the next module (``this packed Modulation Code is what Module~2's parser splits into \texttt{mod\_kind} and \texttt{pkt\_type}'').
\end{itemize}

The transition sentences are the spine of the presentation --- they are what makes five modules feel like one project instead of five.

\subsection{Step 3 --- Live demo preparation}
The live demonstration is the most impactful part of the presentation and also the most failure-prone. Prepare four redundancies:

\begin{tcolorbox}[warning={Live Demo Redundancy Checklist}]
	\begin{enumerate}[leftmargin=1.2em]
		\item \textbf{Pre-recorded backup video.} Record a 3--4 minute screen capture of the inference script running successfully, including the SF reconfiguration moment \emph{and} a clean PAUSE button-hold cycle (press, both Target USART2 streams reacting, release, recovery). If anything fails live (\mbox{USB} port not recognised, Target battery dead, bitstream corrupted), switch to the video without apology and without pausing the presentation.
		\item \textbf{Pre-test in the presentation room.} \mbox{RF} propagation in the presentation space differs from the lab. Run the bidirectional system for 5 minutes in the actual room before your slot, and exercise PAUSE at least twice to confirm the Threat's RSSI table is populated and stable.
		\item \textbf{Minimise configuration changes.} Changing one Target's SF live is impressive; changing multiple parameters simultaneously risks confusion. Pick one SF change plus one PAUSE press, rehearse them, and stick to that script.
		\item \textbf{Start the system during the architecture slides.} The bidirectional ping-pong should be up and running by the time you reach the demo slide. The ``live'' portion of the demo is you switching to the already-running terminals (host inference + both Target USART2 consoles), not you starting something cold.
	\end{enumerate}
\end{tcolorbox}

\begin{tcolorbox}[warning={Rollback Trigger Criteria --- Switch to Backup Video If}]
	\begin{itemize}[leftmargin=1.2em]
		\item \textbf{No classification output for $>$20\,s} after reaching the demo slide (system has stalled).
		\item \textbf{Two consecutive \mbox{CRC} rejects or parser rejects} visible in the terminal (link corruption that will not self-heal during the demo).
		\item \textbf{Reconfigured Target fails to update within 15\,s} of the parameter change (Target firmware crashed or USB reset dropped).
		\item \textbf{PAUSE press produces no \texttt{I'm Jammed} on either Target's USART2 within 1\,s} (Threat RSSI table empty, PAUSE scheduler not running, or the PAUSE button wiring failed).
		\item \textbf{Any \texttt{Traceback} in the inference terminal} (Python-side failure; do not try to recover on stage).
		\item \textbf{Projector or terminal text not visible from the back row} (presentation readability is a hard rollback trigger even if the system is healthy).
	\end{itemize}
	If any trigger fires: say one short line (``Let me cut to the recorded run for time''), switch to the backup video, and continue. Do not debug on stage.
\end{tcolorbox}

\subsection{Step 4 --- Demo script}
Write and rehearse a verbatim script for the live demo. Have three terminals visible: the host four-task inference, Target~A's USART2 console, and Target~B's USART2 console.

\begin{enumerate}[leftmargin=2em]
	\item \emph{``On screen is the live output of our inference script. Each line represents one packet classified on four independent tasks: spreading factor, modulation kind, emitter ID, and packet type.''}
	\item \emph{``You can see Target~A and Target~B in a steady DATA/ACK ping-pong, both at SF7, both LoRa, with the correct emitter IDs (0x01 and 0x02), and the packet-type column alternating DATA/ACK as expected.''}
	\item (Reconfigure Target~A to SF10; wait for the next DATA frame.)
	\item \emph{``I've just reconfigured Target~A to SF10. Watch Target~A's row switch on its next DATA frame --- there --- while Target~B's ACKs remain at SF7.''}
	\item (Return Target~A to SF7. Wait for the link to settle for at least 5\,s.)
	\item \emph{``Now I will hold the PAUSE button on the Threat interceptor. The Threat reads its RSSI table, picks the strongest target, and emits a 32-byte PAUSE row addressed to that ID at a 100\,ms cadence.''}
	\item (Press and hold the user button on the Threat for $\sim$3 seconds.)
	\item \emph{``Watch the targeted Target's USART2 console: there --- it just printed `I'm Jammed' as soon as it decoded the first PAUSE row addressed to its ID. The classifier on the host also tagged that frame as packet-type PAUSE.''}
	\item (Release the button.)
	\item \emph{``The Target waits out its 500\,ms hold-off and the ping-pong resumes --- there. The end-to-end button-to-acknowledgement latency you just saw is roughly the LH5-C S5 number on the slide.''}
\end{enumerate}

Total demo runtime, including narration, should be 3--4 minutes. Any longer and the audience's attention budget for the remaining slides suffers.

\subsection{Step 5 --- Rehearsal}
Two full rehearsals, ideally on different days:

\begin{enumerate}[leftmargin=2em]
	\item \textbf{First rehearsal --- timing.} Present to an empty room or to one peer. Measure slide-by-slide time. If any slide takes more than 90 seconds, cut content. If the total exceeds 20 minutes, cut slides.
	\item \textbf{Second rehearsal --- questions.} Present to a peer or advisor who has not seen the project details. Ask them to interrupt with questions. Note every question --- some will appear again in the actual \mbox{Q\&A} session.
\end{enumerate}

\subsection{Step 6 --- Anticipate Q\&A topics}
The 10-minute \mbox{Q\&A} is often where the final impression is made. Prepare one-sentence answers to at least the following likely questions:

\begin{itemize}[leftmargin=2em]
	\item Why did you choose a random forest (or whichever model won) on the packet-type task? Cite the LH4-H rubric.
	\item What is the operating-condition shift $\Delta_{\text{op-shift}}$ and what causes it? Cite LH4-F + LH5-D B5.
	\item Is the PAUSE protocol FCC-compliant? (Yes: it is a directed protocol message at the same +14\,dBm and 906.5\,MHz as the cooperative DATA/ACK traffic, only emitted when the operator presses the button, with a 100\,ms cadence and 500\,ms hold-off --- not a wide-band jammer.)
	\item If you had to pick one thing to redo, what would it be? Use the LH5-G Discussion answer.
	\item What is the largest cross-domain dependency you found? Cite the LH5-E catalog (F5 or F7 are good answers).
	\item How would this scale to 10 active stations? 100? Discuss FIFO depth, SX1276 half-duplex, and RSSI-table aging.
	\item What is the \FPGA{} resource utilization? Cite Module~2 final utilization report.
	\item Why did you use VS Code with the STMicro extension stack rather than standalone STM32CubeIDE for firmware? Cite the Module~1 reference environment choice.
	\item Why are all three radios on 906.5\,MHz inside the 915\,MHz ISM band? (Operating-point pin-down for the LH5-D regression check; a single carrier eliminates one axis of variability.)
\end{itemize}

\subsection{Step 7 --- Submit the complete package}
Package submission is the last act of the project. Check:
\begin{itemize}[leftmargin=2em]
	\item Final report (from LH5-G) in both PDF and source form.
	\item Presentation slides (PDF + original source).
	\item Backup demo video.
	\item Complete curriculum package (from LH5-F).
	\item All repositories tagged with a final release version.
\end{itemize}

\subsection{Step 8 --- Deliver the presentation}
On the day of the presentation:
\begin{itemize}[leftmargin=2em]
	\item Arrive 30 minutes early. Test the projector connection. Test audio if the backup video has sound.
	\item Start the bidirectional Target~A / Target~B ping-pong and the host inference during your architecture slides (Step~3 redundancy \#4). Confirm the Threat is listening and its RSSI table has at least two entries before reaching the demo slide.
	\item Do not rush transitions. A 2-second silence between slides reads as confidence, not hesitation.
	\item During \mbox{Q\&A}, if a question stumps you, say ``I don't know --- that would be an interesting follow-up experiment''. Do not invent.
\end{itemize}

\begin{tcolorbox}[infobox={The Closing Message}]
	Every presentation about an engineering project is implicitly answering one question: ``What did you learn?'' The instinct is to answer with what was built --- the radios, the \FPGA{} pipeline, the four-task classifier. The better answer is to describe a \emph{cross-module} insight: a case where a decision in one domain constrained or enabled something unexpected in another. That is the answer the framework is designed to produce. End on that insight.
\end{tcolorbox}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] Slide deck built ($\sim$18--20 slides) following the Step~1 outline.
	\item[$\square$] Live demo script written and rehearsed, including the PAUSE button-hold sequence.
	\item[$\square$] Backup demo video recorded (covers SF reconfiguration \emph{and} PAUSE press/release).
	\item[$\square$] Two rehearsals completed; timing fits 20 minutes.
	\item[$\square$] \mbox{Q\&A} prep list of at least 7 likely questions with one-sentence answers, including the FCC-compliance and PAUSE-protocol questions.
	\item[$\square$] Final submission package (report, slides, video, curriculum) assembled and delivered.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-H Deliverable --- Capstone Final Submission}]
	\begin{enumerate}[leftmargin=1.2em]
		\item Presentation slides (PDF + source).
		\item Backup demo video (MP4 or equivalent, 3--4 minutes; covers four-task classification, SF reconfiguration, and PAUSE button-hold).
		\item Delivered presentation (recorded if policy permits, for archival).
		\item Final submission archive bundling: (a) final report, (b) slides, (c) backup video, (d) complete curriculum package, (e) all trained models and datasets.
		\item Post-presentation reflection note (one page): one thing that worked, one thing that did not, one thing you would change if presenting again.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
