\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-F}
\newcommand{\labtitleshort}{Curriculum Package and Instructor Guide}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-F}{Curriculum Package Assembly, Verification, and Instructor's Guide}

\section{Objective}
Assemble the complete, self-contained curriculum package that a future instructor or student can use without contacting the original developer, and write the instructor's guide that shepherds them through hardware setup, software setup, common issues, and curriculum adaptation. This handout produces the primary external artefact of the project.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handout~5-E complete (failure mode catalog available; the ``common issues'' section of the instructor's guide will cite it).
	\item All prior module deliverables present on disk in their final form (lesson plan PDFs, firmware projects, Vivado projects, Python code, trained models, labeled dataset).
\end{itemize}

\section{Procedure}

\subsection{Step 1 --- Verify package contents}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c l X c}
		\toprule
		\textbf{\#} & \textbf{Item} & \textbf{Description} & \textbf{Verified?} \\
		\midrule
		1  & Lesson Plans (Modules 1--5)               & PDF + LaTeX source for each module and all 40 lab handouts; includes curriculum sequence overview & $\square$ \\
		2  & Cooperative Target Firmware (TargetA, TargetB) & Two VS Code STM32 projects; SPI driver, packet builder with packed Modulation Code, FSM with PAUSE listener (ID 0x01 / 0x02) & $\square$ \\
		3  & Threat (Interceptor) Firmware             & VS Code STM32 project; ISR, ring buffer, UART forwarder, RSSI table, button-gated PAUSE TX scheduler (ID 0xFF) & $\square$ \\
		4  & \FPGA{} \mbox{RTL{}}                      & Vivado project; all \mbox{VHDL} sources, XDC, testbenches & $\square$ \\
		5  & Host Collection Script                    & Python \texttt{collect.py} with campaign-config \mbox{JSON}s for C1--C5 & $\square$ \\
		6  & ML Pipeline Code                          & \texttt{preprocess.py}, training scripts (\texttt{baselines.py}, \texttt{svm\_train.py}, \texttt{mlp\_train.py}, \texttt{cnn\_train.py}), \texttt{stratified.py}, \texttt{select\_model.py}, \texttt{inference.py}, \texttt{live\_inference.py} (three deployed tasks: \texttt{sf}, \texttt{beacon}, \texttt{pkt}; \texttt{mod} dropped as single-class) & $\square$ \\
		7  & Trained Models                            & \texttt{.pkl} / \texttt{.pt} files for RF, SVM, MLP, CNN-1D, one set per task; \texttt{deployment\_manifest.json} pins which is deployed per task & $\square$ \\
		8  & Labeled Dataset                           & \texttt{merged\_dataset.csv} + \texttt{data\_dictionary.md} (with \texttt{regime} column) & $\square$ \\
		9  & Hardware \mbox{BOM}                       & Part numbers, vendors, costs (3$\times$ Nucleo, 3$\times$ RFM95W, Nexys A7) & $\square$ \\
		10 & Instructor's Guide                        & The document written in Step~3 of this handout & $\square$ \\
		11 & Package Manifest (\texttt{package\_manifest.json}) & SHA-256 hash + size for every shipped artefact (generated by \texttt{make\_manifest.py} in Step~1) & $\square$ \\
		\bottomrule
	\end{tabularx}
\end{table}

Verification is not just presence --- for each item, confirm it is usable:

\begin{itemize}[leftmargin=2em]
	\item For firmware projects (\#2--\#3): clone to a scratch directory, open in VS Code, and build all three projects (TargetA, TargetB, Threat). No missing \mbox{SDK} references, no unresolved \texttt{\#include}s, no hardcoded paths pointing to the original developer's machine. Confirm each project's \texttt{Beacon\_ID} \texttt{\#define} is set to the correct value (0x01, 0x02, 0xFF).
	\item For the Vivado project (\#4): open in Vivado, run Synthesis. It must complete without warnings about missing files or unconstrained ports beyond those already documented.
	\item For the Python code (\#5, \#6): create a fresh conda or venv environment from a \texttt{requirements.txt}, install, and run the scripts on the packaged dataset. If a script requires a \mbox{USB} device, provide a \texttt{--replay} mode or a dummy input for package-integrity testing.
	\item For the trained models (\#7): \texttt{joblib.load()} / \texttt{torch.load()} succeeds and \texttt{InferenceEngine.predict()} on a sample feature dict returns a sensible class for each of the three deployed tasks (\texttt{sf}, \texttt{beacon}, \texttt{pkt}).
	\item For the dataset (\#8): \texttt{pd.read\_csv()} + the LH4-A thirteen-check audit passes; the \texttt{regime} column contains both ``clean'' and ``contended'' values.
\end{itemize}

\subsubsection*{Package manifest (required for archive integrity)}
Generate a machine-readable \texttt{package\_manifest.json} at the root of the package. It records a SHA-256 hash, size, and version pin for every shipped artefact, so anyone unpacking the archive can verify nothing was silently corrupted or swapped. Run this once at the end of packaging; commit the resulting manifest into the package itself.

\begin{lstlisting}[language=Python, caption={Reference manifest generator (\texttt{make\_manifest.py}).}]
import hashlib, json, os, subprocess, sys
from pathlib import Path

ROOT = Path("capstone_package").resolve()
IGNORE = {"__pycache__", ".git", ".vscode", "build", "dist"}

def sha256(p: Path, buf=1 << 20):
    h = hashlib.sha256()
    with p.open("rb") as f:
        for chunk in iter(lambda: f.read(buf), b""):
            h.update(chunk)
    return h.hexdigest()

def walk(root):
    for dirpath, dirnames, filenames in os.walk(root):
        dirnames[:] = [d for d in dirnames if d not in IGNORE]
        for fn in filenames:
            p = Path(dirpath) / fn
            rel = p.relative_to(root).as_posix()
            yield rel, p

def git_rev():
    try:
        return subprocess.check_output(
            ["git", "rev-parse", "--short=12", "HEAD"],
            cwd=ROOT, text=True).strip()
    except Exception:
        return None

entries = []
for rel, p in walk(ROOT):
    entries.append({
        "path":   rel,
        "bytes":  p.stat().st_size,
        "sha256": sha256(p),
    })

manifest = {
    "package":    "rf-classifier-capstone",
    "git_commit": git_rev(),
    "file_count": len(entries),
    "total_bytes": sum(e["bytes"] for e in entries),
    "files":      sorted(entries, key=lambda e: e["path"]),
}
(ROOT / "package_manifest.json").write_text(
    json.dumps(manifest, indent=2))
print(f"Wrote {ROOT / 'package_manifest.json'} "
      f"({manifest['file_count']} files)")
\end{lstlisting}

\textbf{Verification:} after generating the manifest, re-compute the hash of one large file (e.g.\ \texttt{Data/merged\_dataset.csv}) with an external tool (\texttt{sha256sum}, \texttt{Get-FileHash}, etc.) and confirm it matches the manifest entry. This proves the manifest generator itself is trustworthy before the package ships.

\subsection{Step 2 --- Directory layout}

\begin{lstlisting}[caption={Canonical package layout. Top-level files are the entry points a future user will reach for first.}]
capstone_package/
|-- README.md                       # one-page project overview + entry points
|-- InstructorGuide.pdf             # the document from Step 3 of this handout
|-- LICENSE                         # if applicable
|-- LessonPlans/
|   |-- Module1.pdf / .tex
|   |-- Module2.pdf / .tex
|   |-- Module3.pdf / .tex
|   |-- Module4.pdf / .tex
|   |-- Module5.pdf / .tex
|   |-- glossary.tex
|   |-- Module 1/LabHandouts/       # same structure Modules 1-5
|   |-- ...
|-- Firmware/
|   |-- TargetA/                    # VS Code STM32 project (DATA initiator, ID 0x01)
|   |-- TargetB/                    # VS Code STM32 project (ACK responder, ID 0x02)
|   |-- Threat/                     # VS Code STM32 project (interceptor + PAUSE, ID 0xFF)
|-- FPGA/
|   |-- ingest_top/                 # Vivado project folder
|   |-- bitstreams/                 # ingest_top_debug.bit, ingest_top_prod.bit
|-- Host/
|   |-- collect.py
|   |-- preprocess.py
|   |-- baselines.py / svm_train.py / mlp_train.py / cnn_train.py
|   |-- stratified.py / select_model.py
|   |-- inference.py
|   |-- live_inference.py
|   |-- deployment_manifest.json    # which model is deployed per task
|   |-- requirements.txt
|   |-- campaign_configs/*.json     # C1, C2, C3, C4, C5
|-- Data/
|   |-- merged_dataset.csv          # includes the `regime` column (clean / contended)
|   |-- data_dictionary.md
|   |-- validation_manifest.json
|-- Models/
|   |-- sf/rf.pkl, svm.pkl, mlp.pt, cnn.pt, preprocessor.pkl
|   |-- beacon/...                  # was "id" in earlier drafts
|   |-- pkt/...
|   # Note: no `mod/` subfolder --- single-class skip during M3
|-- Hardware/
|   |-- BOM.md
|   |-- wiring_diagram.png
|   |-- pin_assignments.md
\end{lstlisting}

\subsection{Step 3 --- Write the instructor's guide}
The instructor's guide (\texttt{InstructorGuide.pdf}) has five sections:

\begin{enumerate}[leftmargin=2em]
	\item \textbf{Hardware Setup} (2--3 pages). Assembly instructions for the benchtop platform from the \mbox{BOM}. Include photos of the finished wiring for all three Nucleo+RFM95W stations (Target~A, Target~B, Threat), antenna placement, board-to-board connections, and \mbox{PMOD} pinout. A future student replicating the project should not need to ask ``where does this wire go''. Note explicitly that the PAUSE button on the Threat is the unmodified Nucleo user button (PC13).
	\item \textbf{Software Setup} (2--3 pages). Version-pinned installation instructions for:
	\begin{itemize}[leftmargin=1.2em]
		\item VS Code with the official STMicroelectronics extension stack (Module~1 reference environment; standalone STM32CubeIDE documented as an alternative). Cover the LH1-C ``create all three projects upfront'' procedure, since copying STM32 project folders later breaks the CMake toolchain.
		\item AMD/Xilinx Vivado Design Suite (WebPACK edition, 2024.1 or later).
		\item Python 3.10+ with \texttt{requirements.txt} shipped in the package.
		\item Serial terminal software (any cross-platform option).
	\end{itemize}
	Give exact tested version numbers and note any version mismatches observed during the project.
	\item \textbf{Common Issues} (2--3 pages). A symptom-indexed troubleshooting guide. Draft from the LH5-E failure mode catalog: for each entry, give the symptom as it appears to a new user, then the diagnostic procedure and fix. Examples:
	\begin{itemize}[leftmargin=1.2em]
		\item \textbf{Symptom:} ``\FPGA{} \mbox{LED(15)} stays on after programming.'' Diagnosis: \FIFO{} overflow latched during a startup burst. Fix: reset the board; if it recurs, reduce the Target transmit rate or enlarge the \FIFO{} depth in Module~2.
		\item \textbf{Symptom:} ``Classifier reports SF7 for every packet, even after reconfiguring Target~A.'' Diagnosis: probably label leakage via \texttt{spread\_factor} in the feature set (LH4-B forbidden list regression). Fix: re-run LH4-B with the correct FORBIDDEN map and retrain.
		\item \textbf{Symptom:} ``Holding the Threat's PAUSE button does not stop either Target.'' Diagnosis: RSSI table empty (no DATA frames received yet) or USART2 from Target not wired to the host serial console. Fix: confirm DATA/ACK ping-pong is running first; then re-check the USART2 pin assignment per LH1-F.
		\item \textbf{Symptom:} ``Both Targets print \texttt{I'm Jammed} but the wrong one does so first.'' Diagnosis: RSSI tie at the Threat. Fix: this is the F5/F6 failure mode in LH5-E; document and continue, do not patch the firmware to add hysteresis without also re-running LH5-A.
	\end{itemize}
	\item \textbf{Adaptation Guidance} (1--2 pages). How to scale the curriculum for different contexts:
	\begin{itemize}[leftmargin=1.2em]
		\item Different \FPGA{} board (port \mbox{XDC}, re-check clock constraints).
		\item Different \mbox{ISM} band (sub-GHz 868\,MHz, 433\,MHz regional variants --- update the SX1276 frequency triple and confirm regulatory compliance).
		\item Compressed schedule (which modules can merge; which cannot).
		\item Multi-student version (task allocation across a 3--4 person team --- e.g.\ one student per Target firmware, one for FPGA, one for ML).
	\end{itemize}
	\item \textbf{Assessment Guidance} (1 page). The Module~1--5 rubric with notes on which objectives are essential vs.\ stretch, and suggested grading modifications for senior undergrad vs.\ graduate cohorts.
\end{enumerate}

\subsection{Step 4 --- Write the top-level README}

\begin{lstlisting}[caption={\texttt{README.md} --- first thing any new user sees. Keep it under one page printed.}]
# Capstone: AI-Driven RF Signal Processing

End-to-end teaching framework that builds, from scratch, a system that
classifies Radio-Frequency transmissions in real time on a three-task
suite (spreading factor `sf`, beacon ID `beacon`, packet type `pkt`;
the `mod` task is single-class in this dataset and not deployed):

  Target A (DATA) <-> Target B (ACK)        on 906.5 MHz
            |
            v   (RSSI-targeted PAUSE on operator press)
  Threat interceptor  ->  FPGA pipeline  ->  ML classifier (host)

## Start here
- Read `InstructorGuide.pdf` if you are teaching the course.
- Read `LessonPlans/Module1.pdf` if you are a student beginning the work.
- Read `Hardware/BOM.md` before ordering parts.

## Repo layout
See `InstructorGuide.pdf` section "Package Layout" for the full tree.

## Prerequisites
- STM32 Nucleo-L476RG x3 + RFM95W x3 + 915 MHz ISM-band antennas
  (carrier programmed to 906.5 MHz: SX1276 0xE2 / 0xA0 / 0x00).
- Digilent Nexys A7-100T.
- Host PC with VS Code (STMicro extensions), Vivado 2024.1+, Python 3.10+.

## Verified environment
Windows 11, Vivado 2024.1, VS Code 1.95, Python 3.11. See
`Host/requirements.txt` for pinned Python library versions.

## Current status
All lesson plans + 40 lab handouts + curriculum package completed. Runtime
envelope in `LessonPlans/Module 5/SystemSpecification.pdf`.
\end{lstlisting}

\subsection{Step 5 --- Self-test on a clean machine}
The final verification --- and the one that catches everything the in-situ tests missed --- is to install the package on a \emph{second machine} that was not used during development. If a second machine is unavailable, create a fresh Windows \mbox{VM} or a cleaned Docker container (for the Python portion). Go through the instructor's guide step by step. Every instruction that is unclear, wrong, or missing becomes a change to the guide.

\begin{tcolorbox}[infobox={Why a Clean-Machine Test Matters}]
	In-place testing always overstates portability because the original developer has years of accumulated environment configuration (driver installs, path entries, \mbox{SDK} caches, hardware quirks) that new users do not. The clean-machine test is the only way to find these hidden assumptions. Expect at least 3--5 issues on the first pass; fix them, and run the test again until the guide is followable in one sitting by a user who has never seen the project.
\end{tcolorbox}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] All 11 package items present and usability-verified (items \#2--\#3 cover three distinct firmware projects: TargetA, TargetB, Threat).
	\item[$\square$] Package laid out per the canonical directory tree.
	\item[$\square$] \texttt{InstructorGuide.pdf} written with all five sections.
	\item[$\square$] Top-level \texttt{README.md} concise and accurate.
	\item[$\square$] Clean-machine test completed; issues surfaced and fixed.
	\item[$\square$] Common Issues section cross-references the LH5-E failure catalog (including the F5/F6 PAUSE entries).
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-F Deliverable --- and Session 19 Final Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item Full \texttt{capstone\_package/} directory (or archive) laid out per Step~2.
		\item \texttt{InstructorGuide.pdf}: 8--12 pages, all five sections.
		\item \texttt{README.md} at the package root.
		\item Clean-machine test log: list of issues surfaced, with disposition (fixed / documented / deferred).
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
