\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-D}
\newcommand{\labtitleshort}{Benchmarking Matrix and Spec Table}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-D}{Benchmarking Matrix (B1--B6), Baseline Regression, and System Specification Table}

\section{Objective}
Run the six benchmarking operating points, compare the stressed-system performance against the Module~3 and Module~4 baselines, and fill in the final Module~5 system specification table. The deliverable is a single-page spec sheet that any future user of this system can refer to for operating envelope, three-task accuracy (\texttt{sf}, \texttt{beacon}, \texttt{pkt} --- the \texttt{mod} task is single-class in the dataset and not deployed), expected latency, and PAUSE behavioural-response latency.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handout~5-C complete (per-stage \mbox{GPIO}-toggle instrumentation working at SF7; S5 PAUSE response intervals captured).
	\item Module~3 LH3-D measurements on file (\PER{}, latency at SF7/SF10).
	\item Module~4 LH4-H microbenchmark on file (p99 host inference latency for the three-task suite; deployed mix per \texttt{deployment\_manifest.json} is CNN(\texttt{sf}) + SVM(\texttt{beacon}) + SVM(\texttt{pkt})).
	\item Module~4 LH4-F operating-condition shift results on file ($\Delta_{\text{op-shift}}$ per task).
\end{itemize}

\section{Procedure}

\subsection{Step 1 --- Run the six operating points}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c c c c c X}
		\toprule
		\textbf{ID} & \textbf{SF} & \textbf{Active stations} & \textbf{Distance} & \textbf{Power} & \textbf{Key measurement} \\
		\midrule
		B1 & SF7  & Target~A only                      & 3\,m & +14\,dBm & Minimum latency operating point \\
		B2 & SF10 & Target~A only                      & 3\,m & +14\,dBm & Mid-range latency \\
		B3 & SF12 & Target~A only                      & 3\,m & +14\,dBm & Maximum latency (longest on-air time) \\
		B4 & SF7  & Target~A + Target~B (ping-pong)    & 3\,m & +14\,dBm & Bidirectional DATA/ACK throughput \\
		B5 & SF7  & Target~A + Target~B + Threat PAUSE & 3\,m & +14\,dBm & Contended regime; PAUSE-row classification \\
		B6 & SF7  & Target~A only                      & 8\,m & +2\,dBm  & Low-\mbox{SNR} impact on all stages \\
		\bottomrule
	\end{tabularx}
\end{table}

For each operating point, capture 50 $T_{\text{total}}$ measurements with the LH5-C \mbox{GPIO} setup. Report:
\begin{itemize}[leftmargin=2em]
	\item Mean, std, min, max, and p99 of $T_{\text{total}}$.
	\item Per-stage decomposition (six stages from LH5-C).
	\item The dominant stage (the stage contributing $>$50\% of $T_{\text{total}}$).
	\item Per-task accuracy across all three deployed classifiers (\texttt{sf}, \texttt{beacon}, \texttt{pkt}).
\end{itemize}

\subsection{Step 2 --- Latency histograms}
Plot two latency histograms:
\begin{itemize}[leftmargin=2em]
	\item B1 (lowest latency: SF7, Target~A only). Annotate mean, std, and p99.
	\item B3 (highest latency: SF12, Target~A only). Same annotations.
\end{itemize}

The two should look qualitatively similar (same shape) but translated along the time axis by the $T_{\text{RF}}$ difference ($\sim$1450\,ms between SF7 and SF12 for a 32-byte payload).

\subsection{Step 3 --- Dominance bar chart}
For each of the six operating points, draw a stacked bar showing the proportional contribution of each of the six stages. Observation: on every bar, the \mbox{RF} stage should dominate. This is the \emph{visual} form of the architectural conclusion.

\subsection{Step 4 --- Regression check against Module 3 and Module 4 baselines}

\begin{tcolorbox}[warning={Comparability Rules --- Required Normalization}]
A regression claim is only meaningful if the M3, M4, and M5 numbers were measured at the \emph{same} operating point. Before populating the table, pin down all eight axes below; if any differs across modules, either re-measure the earlier number under the current point or explicitly annotate the divergence in the ``Change'' column.
\begin{enumerate}[leftmargin=1.2em]
	\item \textbf{Spreading factor} (SF7 for the baseline row).
	\item \textbf{Active-station configuration} (e.g.\ Target~A + Target~B ping-pong, PAUSE button not held).
	\item \textbf{Distance} (3\,m for the baseline row).
	\item \textbf{TX power} (+14\,dBm for the baseline row).
	\item \textbf{Carrier frequency} (906.5\,MHz; record the SX1276 register triple 0xE2/0xA0/0x00).
	\item \textbf{Firmware commit hashes} (Target~A + Target~B + Threat builds) --- record the short SHA used in each module.
	\item \textbf{Bitstream hash} (the \texttt{.bit} file hash; differs between debug and production builds).
	\item \textbf{Model artefact versions} (the three deployed-model paths + content hashes used for the M4 and M5 accuracy numbers, one per task --- file extensions vary: \texttt{cnn.pt} for \texttt{sf}, \texttt{svm.pkl} for \texttt{beacon} and \texttt{pkt} per the current \texttt{deployment\_manifest.json}).
\end{enumerate}
Record these in the row header of the table below or in an ``Operating Point'' cell. Any row without all eight fields pinned is not a regression check; it is an exploration.
\end{tcolorbox}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l c c c X}
		\toprule
		\textbf{Metric} & \textbf{M3} & \textbf{M4} & \textbf{M5} & \textbf{Change / root cause if regressed} \\
		\midrule
		\PER{} (A+B ping-pong, SF7)                & & & & \\
		End-to-end latency (SF7, 3\,m)             & & --- & & \\
		\texttt{sf} classification accuracy        & --- & & & \\
		\texttt{beacon} (emitter ID) accuracy      & --- & & & \\
		\texttt{pkt} (DATA/ACK/PAUSE) accuracy     & --- & & & \\
		$\Delta_{\text{op-shift}}$ (clean$\to$contended) & --- & & & \\
		\FIFO{} peak utilization (A+B)             & & --- & & \\
		\bottomrule
	\end{tabularx}
\end{table}

Any metric that has drifted $>$5\% from the Module~3 or Module~4 number at the \emph{same} operating point (all eight axes matched) is a regression. Before concluding the system degraded, rule out:

\begin{tcolorbox}[warning={Regression Investigation Order}]
	\begin{enumerate}[leftmargin=1.2em]
		\item \textbf{Hardware:} antenna loose, USB cable partially disconnected, breadboard contact oxidation, Target battery low. Re-run the LH3-A pre-integration checklist.
		\item \textbf{Firmware:} Target~A, Target~B, or Threat firmware modified during Module~4 development (e.g., baud rate changed for debugging and not reverted, or PAUSE hold-off duration tweaked).
		\item \textbf{Bitstream:} \FPGA{} bitstream regenerated with \mbox{ILA{}} probes added/removed, altering timing or feature-vector packing.
		\item \textbf{Environment:} new equipment in the lab radiating near 906.5\,MHz, furniture moved, obstructions introduced between LH3 and LH5.
	\end{enumerate}
	Only after all four rule-outs fail should you conclude the system itself degraded. Document the investigation in the deliverable regardless of outcome --- the \emph{absence} of a regression is also a valuable result, proving the system is stable across modules.
\end{tcolorbox}

\subsection{Step 5 --- Fill the system specification table}
This is the headline Module~5 artefact:

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{X c l}
		\toprule
		\textbf{Parameter} & \textbf{Value} & \textbf{Source} \\
		\midrule
		Maximum concurrent active stations (\PER{} $<$ 5\%)         & \rule{1.5cm}{0.4pt}    & LH5-A (S1) \\
		Maximum frame rate (\FIFO{} $<$ 75\%)                       & \rule{1.5cm}{0.4pt}\,fps & LH5-A (S1) \\
		Minimum classifiable \mbox{SNR} (per-task accuracy $>$ 80\%) & \rule{1.5cm}{0.4pt}\,dB & LH5-B (S3) \\
		Maximum classifiable distance (per-task accuracy $>$ 80\%)   & \rule{1.5cm}{0.4pt}\,m  & LH5-B (S3) \\
		End-to-end latency at SF7 (mean $\pm$ std)                  & \rule{1.5cm}{0.4pt}\,ms & B1 \\
		End-to-end latency at SF12 (mean $\pm$ std)                 & \rule{1.5cm}{0.4pt}\,ms & B3 \\
		\texttt{sf} classification accuracy (nominal)               & \rule{1.5cm}{0.4pt}\%   & B1 \\
		\texttt{beacon} (emitter ID) accuracy (nominal)             & \rule{1.5cm}{0.4pt}\%   & B1 \\
		\texttt{pkt} (DATA/ACK/PAUSE) accuracy (nominal)            & \rule{1.5cm}{0.4pt}\%   & B5 \\
		Operating-condition shift $\Delta_{\text{op-shift}}$ (worst task) & \rule{1.5cm}{0.4pt}\,pp & LH4-F + B5 \\
		PAUSE response latency $T_{\text{btn}\to\text{ACK}}$ (p99)  & \rule{1.5cm}{0.4pt}\,ms & LH5-C (S5) \\
		PAUSE recovery latency $T_{\text{recover}}$ (p99)           & \rule{1.5cm}{0.4pt}\,ms & LH5-C (S5) \\
		RSSI-targeting accuracy (correct Target~ID selected)        & \rule{1.5cm}{0.4pt}\%   & LH5-C (S5) \\
		Maximum stable run time (tested)                            & \rule{1.5cm}{0.4pt}\,min & LH5-B (S4) \\
		\FPGA{} resource utilization (\mbox{LUT{}} / FF / \mbox{BRAM{}}) & \rule{1.5cm}{0.4pt}  & Module~2 \\
		Host inference latency (three-task, p99)                    & \rule{1.5cm}{0.4pt}\,ms & Module~4 \\
		\bottomrule
	\end{tabularx}
	\caption{Module~5 system specification table. Every value is cited back to its source experiment; no assumed or theoretical numbers.}
\end{table}

\subsection{Step 6 --- Architectural implications paragraph}
Write a paragraph (150--250 words) addressing:

\begin{enumerate}[leftmargin=2em]
	\item Which stage dominates $T_{\text{total}}$ at every operating point? (Spoiler: \mbox{RF}.)
	\item Given this, would moving the three-task Module~4 inference step onto the \FPGA{} materially reduce latency? Quantify your answer using the LH5-D bar chart proportions.
	\item Under what deployment scenario \emph{would} on-\FPGA{} inference be worthwhile? (The honest answer is: only when the host \mbox{PC} is eliminated entirely, for standalone embedded deployment.)
	\item Given the \mbox{RF}-dominated architecture, what is the single highest-leverage change for reducing latency? (Usually: smaller payload or lower SF --- both reduce the on-air time, at the cost of payload capacity or range.)
	\item Does the PAUSE behavioural-response latency (LH5-C S5) sit on the same critical path as $T_{\text{total}}$, or does it have its own bottleneck (the 100\,ms PAUSE cadence)? Explain which spec changes would shorten each.
\end{enumerate}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] All six operating points measured with $\geq$50 samples each.
	\item[$\square$] B1 and B3 histograms generated.
	\item[$\square$] Per-stage dominance bar chart generated for all six points.
	\item[$\square$] Three-task accuracy reported per operating point (B5 is the only point at which packet-type PAUSE rows appear).
	\item[$\square$] Baseline regression table filled; any regressions investigated using the 4-step protocol.
	\item[$\square$] System specification table filled with every value cited.
	\item[$\square$] Architectural-implications paragraph written.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-D Deliverable --- and Session 18 Final Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item B1--B6 measurement table (one row per operating point, columns for mean / std / p99 / dominant stage / per-task accuracy across the three deployed tasks).
		\item Two latency histograms (B1 and B3) with mean / std / p99 annotations.
		\item Stacked dominance bar chart for all six operating points.
		\item Baseline regression table with investigation notes for any regression.
		\item Completed system specification table including PAUSE response/recovery latency and RSSI-targeting accuracy.
		\item Architectural-implications paragraph answering the five questions from Step~6.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
