\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{5-B}
\newcommand{\labtitleshort}{Rapid Change, Low SNR, Endurance (S2--S4)}
\input{LabHandout_Preamble}

\begin{document}
\labheader{5-B}{S2 Rapid Parameter Change, S3 Low-SNR, and S4 Extended-Duration Stress Tests}

\section{Objective}
Execute three of the remaining stress-test dimensions: S2 evaluates whether the deployed three-task classifier suite (\texttt{sf}, \texttt{beacon}, \texttt{pkt}) tracks dynamic SF changes from a cooperative target while a second target remains fixed; S3 extends the accuracy-vs.-\mbox{SNR} curve from Module~4 into the low-\mbox{SNR} regime --- including an intentionally out-of-distribution obstruction test --- and S4 runs the bidirectional system for 30 minutes to surface time-dependent failure modes. The \texttt{mod} task is single-class in the dataset and was dropped during Module~3, so this handout only ever evaluates three tasks. The PAUSE behavioural-response stress test (S5) is covered in LH5-C alongside the latency instrumentation it shares.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handout~5-A complete (stress test plan written; S1a--S1e executed; \texttt{s1\_results.json} in hand with three-task (\texttt{sf}, \texttt{beacon}, \texttt{pkt}) accuracy columns).
	\item Target~A (ID 0x01), Target~B (ID 0x02), and the Threat interceptor (ID 0xFF) all operational on 906.5\,MHz.
	\item Firmware build environment available for in-situ SF reconfiguration on Target~A.
	\item A physical obstruction available for S3e (metal filing cabinet, stack of textbooks, or similar).
\end{itemize}

\section{Procedure}

\subsection{Step 1 --- S2: Rapid Parameter Change}
Configure both cooperative targets at SF7, +14\,dBm, fixed 2-second DATA/ACK ping-pong (Target~A initiates DATA, Target~B replies ACK; PAUSE button \emph{not} held). Start \texttt{live\_inference.py} and confirm the three-task classifier reports SF7, the correct beacon ID per packet, and the correct packet type (DATA vs.\ ACK) for both targets. Then cycle Target~A's SF while Target~B remains at SF7:

\begin{enumerate}[leftmargin=2em]
	\item $t = 0$: Both targets at SF7. Verify inference reports SF7, correct emitter ID, and correct packet type for every frame.
	\item $t = 30$\,s: Reflash or reconfigure Target~A to SF10. Observe the number of packets between reflash completion and the first correct SF10 classification on Target~A's DATA frames. Record whether Target~B's SF7 ACK frames remain correctly classified throughout (the \texttt{beacon} and \texttt{pkt} tasks should be unaffected; only \texttt{sf} should transition).
	\item $t = 60$\,s: Change Target~A to SF12. Record the same observations.
	\item $t = 90$\,s: Return Target~A to SF7. Record the return behaviour.
	\item Continue for the remainder of 10 minutes, cycling the same sequence every 90 seconds to gather multiple transitions per SF pair.
\end{enumerate}

\begin{tcolorbox}[warning={Transient Misclassification at SF Boundaries}]
	The Module~4 candidate models classify each feature vector independently (the deployed CNN runs over an 8-frame sequence buffer but has no per-Target state) --- there is no ``tracking delay'' inside the model. However, the \FPGA{}'s \texttt{inter\_arrival\_ms} feature is computed relative to the \emph{previous} observation from the same Beacon~ID. On the first packet after a transition, \texttt{inter\_arrival\_ms} is computed against the last old-SF DATA frame from Target~A, whose on-air time is different. This can cause one or two transient misclassifications right at the boundary on the \texttt{sf} task. The \texttt{pkt} and \texttt{beacon} tasks should remain stable throughout because they read directly from the packed Modulation Code byte and the source-ID byte. Document the transients and explain them in terms of the Module~2 feature-computation logic.
\end{tcolorbox}

\subsection{Step 2 --- S2 analysis}
Build a transition table that tracks all three deployed classification tasks separately:

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c c c c c X}
		\toprule
		\textbf{Trans.} & \textbf{From} & \textbf{To} & \textbf{Pkts to first correct SF} & \textbf{Pkt-type / beacon stable?} & \textbf{Notes} \\
		\midrule
		1 & SF7  & SF10 & & & \\
		2 & SF10 & SF12 & & & \\
		3 & SF12 & SF7  & & & \\
		\bottomrule
	\end{tabularx}
\end{table}

Expected: the SF column should be 0, 1, or 2 --- never larger. Packet type and beacon ID should be stable (zero transition packets) because both are decoded directly from the Modulation Code byte and the source-ID byte rather than from inter-arrival statistics. If you consistently see 3+ on the SF column, revisit the \FPGA{} feature computation: is \texttt{inter\_arrival\_ms} being reset on \texttt{beacon\_id} change, or is it measuring across mismatched SF conditions for more than one packet? If packet type or beacon ID ever flips, you have a Modulation Code packing bug and must revisit LH2-D.

\subsection{Step 3 --- S3: Low-SNR operating conditions}

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{c c c c X}
		\toprule
		\textbf{Config} & \textbf{TX Power} & \textbf{Distance} & \textbf{Expected SNR} & \textbf{Purpose} \\
		\midrule
		S3a & +20\,dBm & 3\,m           & High ($>$10\,dB)  & Comfortable baseline \\
		S3b & +14\,dBm & 5\,m           & Moderate (5--10\,dB) & Nominal operating point \\
		S3c & +8\,dBm  & 5\,m           & Low (0--5\,dB)    & Beginning of degradation \\
		S3d & +2\,dBm  & 8\,m           & Very low ($<$0\,dB) & Near interceptor sensitivity \\
		S3e & +2\,dBm  & 8\,m + obstruction & Minimal       & Extreme: partial blockage, out-of-distribution \\
		\bottomrule
	\end{tabularx}
	\caption{S3 low-\mbox{SNR} sweep. Run each configuration for 3 minutes with Target~A only, SF7, fixed 1\,s DATA interval, PAUSE button not held. Target~B and the Threat are powered but not transmitting so the only on-air emitter is Target~A.}
\end{table}

For S3e, place the obstruction physically between Target~A and the Threat interceptor. The resulting multipath and shadowing produce \mbox{RSSI}/\mbox{SNR} samples that fall \emph{outside} the Module~3 training data distribution (the operating-condition shift framing from LH4-F is directly relevant here).

\subsection{Step 4 --- S3 degradation curve}
Plot:
\begin{itemize}[leftmargin=2em]
	\item Measured mean \mbox{RSSI} vs.\ configuration (bar or line chart).
	\item Classification accuracy vs.\ configuration, one line per deployed task --- \emph{three lines total}: \texttt{sf}, \texttt{beacon}, \texttt{pkt}.
	\item Overlay: the Module~4 LH4-G accuracy-vs.-\mbox{SNR} curve (also three lines), so the Module~5 data extends the existing plot into the low-\mbox{SNR} regime rather than duplicating it.
\end{itemize}

Then characterise the degradation mode for each task:
\begin{itemize}[leftmargin=2em]
	\item \textbf{Graceful:} Accuracy decreases smoothly; SF misclassifications favour adjacent classes; packet-type and beacon-ID degrade only when CRC starts failing.
	\item \textbf{Catastrophic:} Accuracy drops abruptly to near-random; misclassifications scatter across all classes.
\end{itemize}

Expected pattern: packet type and beacon ID should hold near 100\% as long as CRC passes (these are byte reads, not statistical inferences) and then collapse abruptly once \texttt{packet\_crc\_ok} starts dropping. The \texttt{sf} task should degrade more gradually because it depends on continuous-valued features (\texttt{rssi\_dbm}, \texttt{snr\_db}, \texttt{inter\_arrival\_ms}). Either outcome is a valid experimental result; the analysis value is in explaining \emph{which} outcome occurred per task and tracing it to the training data distribution (Module~3) and feature-engineering decisions (Module~4).

\subsection{Step 5 --- S4: Extended-duration stability run}
Configure Target~A and Target~B at SF7, +14\,dBm, 2-second fixed DATA/ACK ping-pong (the Module~3 nominal C2 configuration). Threat is powered, listening, and \emph{not} transmitting (PAUSE button not held). Start \texttt{live\_inference.py} with logging enabled and let the system run for 30 minutes uninterrupted.

Monitor and record:

\begin{table}[h!]
	\centering
	\renewcommand{\arraystretch}{1.3}
	\begin{tabularx}{\textwidth}{l X}
		\toprule
		\textbf{Metric} & \textbf{What to watch for} \\
		\midrule
		\texttt{obs\_count} trajectory       & Linear growth; any reset or plateau is abnormal \\
		Sequence-number gap (per beacon ID) & Constant gap = one-time desync; growing gap = ongoing loss; check both 0x01 and 0x02 \\
		DATA / ACK ratio                     & Should sit at 1:1 within $\pm$5\%; a drift indicates one direction of the link is degrading \\
		\FIFO{} baseline utilization         & Slow upward creep = throughput deficit \\
		\FIFO{} overflow alarm (\texttt{led(15)}) & Should never set under nominal load \\
		Per-task accuracy per 5-min window   & All three deployed tasks should be stable; drift suggests feature drift \\
		Python inference-script memory        & Should plateau after $\sim$1 minute; linear growth = leak \\
		Python inference-script \mbox{CPU} usage     & Should be consistent; spikes correlate with \mbox{OS} scheduling \\
		\bottomrule
	\end{tabularx}
\end{table}

\subsection{Step 6 --- S4 post-run analysis}

\begin{tcolorbox}[infobox={CSV logging is a local enhancement to \texttt{live\_inference.py}}]
The shipped \texttt{live\_inference.py} prints to stdout only. For S4 you have two options: (1) \texttt{tee} stdout to a file and parse it with the regex skeleton below, or (2) extend \texttt{live\_inference.py} with a small CSV writer that emits one row per frame containing \texttt{timestamp\_host, beacon\_id, seq, pred\_sf, pred\_beacon, pred\_pkt}. Option 2 is the cleaner path; option 1 is faster.
\end{tcolorbox}

\begin{lstlisting}[language=Python, caption={\texttt{analyse\_s4.py} --- minute-by-minute three-task stability metrics. Assumes you have written a CSV with the columns above (or adapt the parser to whatever your extended \texttt{live\_inference.py} emits). Ground-truth columns (\texttt{gt\_sf}, \texttt{gt\_beacon}, \texttt{gt\_pkt}) come from the known operator configuration during S4 (SF7, beacon 0x01/0x02 by Target, alternating DATA/ACK).}]
import pandas as pd, matplotlib.pyplot as plt
df = pd.read_csv("s4_log.csv")   # produced by your extended live_inference.py
df["ts"] = pd.to_datetime(df["timestamp_host"])
df["minute"] = (df["ts"] - df["ts"].min()).dt.total_seconds() // 60

# Per-minute accuracy across all three deployed tasks
for task in ["sf", "beacon", "pkt"]:
    acc = df.groupby("minute").apply(
        lambda g: (g[f"pred_{task}"] == g[f"gt_{task}"]).mean())
    plt.plot(acc.index, acc.values, label=task)
plt.xlabel("Minute")
plt.ylabel("Accuracy")
plt.title("S4 --- per-minute three-task accuracy (30 min run)")
plt.legend()
plt.tight_layout()
plt.savefig("figs/s4_accuracy_timeseries.png", dpi=150)

# Sequence-number gap per beacon ID (0x01 and 0x02)
gap = df.groupby("beacon_id").apply(
    lambda g: (g["seq"].max() - g["seq"].min() + 1) - len(g))
print("Sequence gap per beacon:")
print(gap)

# DATA/ACK balance
counts = df["pred_pkt"].value_counts()
print("Packet-type counts:", counts.to_dict())
print(f"DATA/ACK ratio: {counts.get('DATA',0) / max(counts.get('ACK',1),1):.3f}")
\end{lstlisting}

\begin{tcolorbox}[infobox={What S4 Typically Surfaces}]
	The 30-minute run is usually clean, but when it is not, the failures are pedagogically valuable:
	\begin{itemize}[leftmargin=1.2em]
		\item \textbf{Memory leak in the Python logger:} often traced to an unbounded list appended once per frame --- a Module~4 code-hygiene issue surfacing only over time.
		\item \textbf{Slow \FIFO{} creep:} sometimes traced to an occasional extra parser stall during \mbox{UART} framing-error recovery --- a Module~2 issue surfacing only under multi-emitter scheduling.
		\item \textbf{Accuracy drift on SF / modulation:} most often traced to thermal \mbox{RSSI} drift in the Threat interceptor's SX1276, which shifts the feature distribution away from training --- a Module~1 hardware phenomenon surfacing only in the Module~4 classifier.
		\item \textbf{DATA/ACK ratio drift:} typically a directional link asymmetry --- one target's antenna orientation or cable seating has changed; check both Target~A and Target~B before blaming the model.
	\end{itemize}
	Each of these is a Module~5 observation with a non-Module~5 root cause. Capture them for the failure-mode catalog in LH5-E.
\end{tcolorbox}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] S2: all three SF transitions executed; three-task transition table filled.
	\item[$\square$] S2: packet-type and beacon-ID columns confirm zero-packet stability across SF transitions.
	\item[$\square$] S3: all five configurations run; per-task degradation mode labelled (graceful vs.\ catastrophic).
	\item[$\square$] S3 plot overlays the LH4-G accuracy-vs.-SNR curve and contains three lines per panel (\texttt{sf}, \texttt{beacon}, \texttt{pkt}).
	\item[$\square$] S4: 30-minute bidirectional run completed without manual intervention; PAUSE never asserted.
	\item[$\square$] S4 per-minute three-task accuracy time-series saved.
	\item[$\square$] S4 DATA/ACK ratio recorded.
	\item[$\square$] Any anomalies from S2/S3/S4 tagged for the LH5-E failure catalog.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 5-B Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item S2 inference log + three-task transition table with the ``packets until first correct'' column filled for \texttt{sf}, and the stability columns confirmed for \texttt{pkt} and \texttt{beacon}.
		\item S3 accuracy-vs.-configuration plot (three task lines) overlaid on the LH4-G Module~4 curve; short paragraph classifying degradation mode per task.
		\item S4 log file (\texttt{s4\_log.csv}), 30-minute time-series plots (three-task accuracy, \FIFO{} peak, DATA/ACK ratio, Python memory), and a one-paragraph stability verdict.
		\item Updated Session-17 summary table: maximum safe active-station count (from LH5-A), minimum classifiable \mbox{SNR} per task, SF-transition latency, 30-min stability verdict.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
