\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{3-H}
\newcommand{\labtitleshort}{Dataset Merging and EDA}
\input{LabHandout_Preamble}

\begin{document}
\labheader{3-H}{Dataset Merging, Exploratory Data Analysis, and Module~4 Handoff}

\section{Objective}
Merge all campaign \mbox{CSV} files into a single, clean, labeled dataset; perform the five required exploratory data analysis (\mbox{EDA}) visualizations; document the operating-condition shift between the clean regime (C1, C2, and C4 single-station DATA-only) and the contended regime (C5 multi-station ping-pong), entirely on hardware data; and produce the data dictionary that serves as the Module~4 data contract. The merged dataset is the primary deliverable of Module~3.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handout~3-G complete (campaigns C1, C2, C4, and C5 executed and validated).
	\item Python with \texttt{pandas}, \texttt{matplotlib}, \texttt{seaborn}, and \texttt{scipy} installed.
\end{itemize}

\section{Procedure}

\subsection{Step 1 --- Merging script}
Place \texttt{merge\_dataset.py} in the \texttt{Code/Tools/CampaignCollect/} folder alongside \texttt{collect.py}. The script auto-discovers every \texttt{results\_*.csv} under the \texttt{logs/} subdirectory, so no file list needs to be maintained as new campaigns are added. Run it with \texttt{python merge\_dataset.py}; it writes \texttt{merged\_dataset.csv} in the same folder.

\begin{lstlisting}[language=Python, caption={Dataset merging script (\texttt{merge\_dataset.py}).}]
import csv
import sys
import random
from pathlib import Path
from collections import Counter, defaultdict

SCRIPT_DIR = Path(__file__).parent
LOGS_DIR   = SCRIPT_DIR / "logs"
OUT_CSV    = SCRIPT_DIR / "merged_dataset.csv"

EXPECTED_COLS = [
    "campaign_id", "run_id", "timestamp_host",
    "rx_timestamp_ms", "tx_timestamp_ms", "inter_arrival_ms",
    "beacon_id", "seq_num", "pkt_type", "mod_kind", "spread_factor",
    "rssi_dbm", "snr_db", "freq_error_hz", "payload_len",
    "distance_m", "label_sf", "label_mod", "mod_kind_label",
    "label_beacon", "label_pkt",
]

SKIP_DIRS  = {"dry_run"}
SKIP_FILES = {"C1_combined.csv", "C2_combined.csv", "merged_dataset.csv"}


def discover_csvs():
    found = []
    for p in sorted(LOGS_DIR.rglob("results_*.csv")):
        if p.parent.name in SKIP_DIRS:
            continue
        if p.name in SKIP_FILES:
            continue
        found.append(p)
    return found


def stratified_split(rows, label_col, fracs=(0.70, 0.15, 0.15), seed=42):
    # Split at run_id level stratified by label_col to prevent leakage:
    # rows within one run share channel state, so splitting rows randomly
    # would leak that context into the test set.
    rng = random.Random(seed)
    by_label = defaultdict(lambda: defaultdict(list))
    for i, row in enumerate(rows):
        by_label[row[label_col]][row["run_id"]].append(i)

    split_map = {}
    for label_val, run_dict in by_label.items():
        runs = sorted(run_dict.keys())
        rng.shuffle(runs)
        n = len(runs)
        n_train = max(1, round(n * fracs[0]))
        n_val   = max(1, round(n * fracs[1]))
        for j, run_id in enumerate(runs):
            tag = ("train" if j < n_train
                   else "val" if j < n_train + n_val
                   else "test")
            for idx in run_dict[run_id]:
                split_map[idx] = tag
    return split_map


csvs = discover_csvs()
if not csvs:
    print(f"No results_*.csv files found under {LOGS_DIR}")
    sys.exit(1)

print(f"Found {len(csvs)} campaign CSV(s):\n")
all_rows = []
for p in csvs:
    with open(p, newline="") as f:
        rows = list(csv.DictReader(f))
    if not rows:
        print(f"  [SKIP] {p.name} - empty")
        continue
    missing = set(EXPECTED_COLS) - set(rows[0].keys())
    if missing:
        print(f"  [WARN] {p.name} missing columns: {missing}")
    cid = rows[0].get("campaign_id", "?")
    rid = rows[0].get("run_id", "?")
    print(f"  {p.name}  ->  campaign={cid}  run={rid}  rows={len(rows)}")
    all_rows.extend(rows)

print(f"\nTotal rows before dedup: {len(all_rows)}")

seen = set()
deduped = []
for row in all_rows:
    key = (row.get("run_id"), row.get("beacon_id"),
           row.get("seq_num"), row.get("pkt_type"))
    if key not in seen:
        seen.add(key)
        deduped.append(row)
print(f"Total rows after dedup:  {len(deduped)}")

for row in deduped:
    row["regime"] = (
        "contended" if row.get("campaign_id", "").startswith("C5")
        else "clean")

split_map = stratified_split(deduped, "label_sf")
for i, row in enumerate(deduped):
    row["split"] = split_map.get(i, "train")

by_campaign = Counter(r["campaign_id"] for r in deduped)
by_regime   = Counter(r["regime"]      for r in deduped)
by_pkt      = Counter(r["pkt_type"]    for r in deduped)
by_split    = Counter(r["split"]       for r in deduped)

print("\n--- Summary ---")
print("Rows per campaign_id:")
for k, v in sorted(by_campaign.items()):
    print(f"  {k}: {v}")
print(f"Regime:   {dict(by_regime)}")
print(f"pkt_type: {dict(by_pkt)}")
print(f"Split:    {dict(by_split)}")

out_cols = EXPECTED_COLS + ["regime", "split"]
with open(OUT_CSV, "w", newline="") as f:
    writer = csv.DictWriter(f, fieldnames=out_cols, extrasaction="ignore")
    writer.writeheader()
    writer.writerows(deduped)

print(f"\nSaved: {OUT_CSV}  ({len(deduped)} rows)")
\end{lstlisting}

\begin{tcolorbox}[warning={Stratification at the run\_id Level}]
	Split at the \emph{run\_id} level, not at \texttt{campaign\_id} or at the individual-sample level. Samples inside one run share channel conditions and hardware state; splitting samples randomly leaks channel context into the test set. Splitting on \texttt{campaign\_id} alone is also unsafe: one campaign family can contain several distinct collection runs. Keying on \texttt{run\_id} (one value per invocation of \texttt{collect.py}) is the only way to keep splits genuinely independent.
\end{tcolorbox}

\begin{tcolorbox}[infobox={Canonical Modulation Kind and Packet Type Mappings}]
Modulation kind is represented throughout Modules~1--5 using the integer codes \textbf{1=LoRa, 2=FSK, 3=OOK} (Module~1 Table~4). The \texttt{mod\_kind} feature vector byte, the \texttt{mod\_kind\_label} config column, and any Module~4 integer checks use this mapping. The string alias \texttt{label\_mod} (``lora''/``fsk''/``ook'') is retained only for readability and never used in a numeric check. Do not introduce a 0/1 alias.

Packet type is \textbf{0=DATA, 1=ACK, 2=PAUSE}. \texttt{pkt\_type} is extracted from the low 2 bits of the radio packet's \texttt{Modulation Code} byte by the LH2-F bit-field extractor and is the canonical authority for the \texttt{label\_pkt} ground-truth label.
\end{tcolorbox}

\newpage
\subsection{Step 2 --- Verify merged dataset statistics}
After running \texttt{merge\_dataset.py}, confirm:
\begin{itemize}[leftmargin=2em]
	\item Total rows $\geq$ 2{,}100 (C1: 800, C2: 800, C4: 800, C5: $\sim$600 = 3000 theoretical; expect some \PER{}-driven shortfall and C4 may be partial depending on execution).
	\item Rows per campaign ID roughly as expected from the target packet counts.
	\item \texttt{regime} column shows roughly 75\% \texttt{clean} (C1, C2, C4) and 25\% \texttt{contended} (C5); the LH4-F operating-condition shift experiment relies on having both regimes well-represented.
	\item \texttt{pkt\_type} value counts: \texttt{0} (DATA) dominates due to C1--C4; \texttt{1} (ACK) appears only in C5b.
	\item \texttt{split} column contains ``train'', ``val'', and ``test'' at approximately 70/15/15\%.
	\item No \texttt{[WARN]} lines in the script output indicating missing schema columns. The script does not compute a \texttt{label\_mismatch} column; verify firmware config correctness by checking that the \texttt{spread\_factor} column matches \texttt{label\_sf} for all DATA rows (\texttt{pkt\_type = 0}).
\end{itemize}

\subsection{Step 3 --- EDA Visualization 1: RSSI by TX power (C1)}
\begin{lstlisting}[language=Python]
import matplotlib.pyplot as plt
import seaborn as sns
import pandas as pd

df = pd.read_csv("merged_dataset.csv")
c1 = df[df["campaign_id"] == "C1"]
# sweep_value (TX power) is not stored as a column; use run_id to recover it.
c1 = c1.copy()
c1["tx_power"] = c1["run_id"].str.extract(r"pwr(\d+)").astype(int)
sns.violinplot(data=c1, x="tx_power", y="rssi_dbm")
plt.xlabel("TX Power (dBm)")
plt.ylabel("RSSI (dBm)")
plt.title("Campaign C1: RSSI Distribution by TX Power")
plt.savefig("eda1_rssi_by_txpower.png", dpi=150)
\end{lstlisting}
Expected: distributions shift right (higher RSSI, less negative) as TX power increases. The distributions should overlap slightly at adjacent power levels, indicating the system is sensitive to 6\,dB power steps but not perfectly separable by power alone.

\subsection{Step 4 --- EDA Visualization 2: RSSI vs.\ distance with FSPL overlay (C4)}
\begin{lstlisting}[language=Python]
import numpy as np

c4 = df[df["campaign_id"] == "C4"]
means = c4.groupby("distance_m")["rssi_dbm"].mean()
stds  = c4.groupby("distance_m")["rssi_dbm"].std()

# FSPL model: RSSI(d) = P_tx - 20*log10(d) - 20*log10(f_Hz) + 147.55
f_hz  = 906.5e6
p_tx  = 14  # dBm
d_range = np.linspace(0.5, 10, 200)
fspl = p_tx - 20*np.log10(d_range) - 20*np.log10(f_hz) + 147.55

plt.figure()
plt.errorbar(means.index, means.values, yerr=stds.values,
             fmt="o", capsize=5, label="Measured mean RSSI")
plt.plot(d_range, fspl, "--", label="FSPL model (+14 dBm, 906.5 MHz)")
plt.xlabel("Distance (m)")
plt.ylabel("RSSI (dBm)")
plt.title("Campaign C4: RSSI vs. Distance")
plt.legend()
plt.savefig("eda2_rssi_vs_distance.png", dpi=150)
\end{lstlisting}

\subsection{Step 5 --- EDA Visualization 3: Feature correlation matrix}
\begin{lstlisting}[language=Python]
numeric_cols = ["rssi_dbm", "snr_db", "inter_arrival_ms",
                "freq_error_hz", "spread_factor", "distance_m"]
corr = df[numeric_cols].astype(float).corr()
sns.heatmap(corr, annot=True, fmt=".2f", cmap="coolwarm", center=0)
plt.title("Feature Correlation Matrix (Full Dataset)")
plt.tight_layout()
plt.savefig("eda3_correlation_matrix.png", dpi=150)
\end{lstlisting}
Identify and discuss in your report: which feature pairs are strongly correlated ($|r| > 0.7$), and whether any correlations might cause problems for a linear classifier.

\newpage
\subsection{Step 6 --- EDA Visualization 4: Inter-arrival time by station count (C5a vs.\ C5b)}
\begin{lstlisting}[language=Python]
c5a = df[df["run_id"].str.startswith("C5a")]["inter_arrival_ms"]
c5b = df[df["run_id"].str.startswith("C5b")]["inter_arrival_ms"]
plt.figure()
plt.hist(c5a, bins=40, alpha=0.6, label="C5a Single station (DATA-only)")
plt.hist(c5b, bins=40, alpha=0.6, label="C5b Ping-pong (DATA + ACK)")
plt.xlabel("Inter-arrival Time (ms)")
plt.ylabel("Count")
plt.title("Inter-arrival Time: Single Station vs. Ping-pong")
plt.legend()
plt.savefig("eda4_interarrival_scheduling.png", dpi=150)
\end{lstlisting}
Expected: C5a produces a tight peak near 2\,s (fixed initiator interval); C5b produces a bimodal distribution --- a DATA peak near 1\,s and a tighter ACK peak near the round-trip latency. The ACK cluster is a learnable temporal signature for emitter identification in Module~4.

\subsection{Step 7 --- EDA Visualization 5: Class balance (4 tasks)}
\begin{lstlisting}[language=Python]
PKT_LABELS = {0: "DATA", 1: "ACK"}
TASKS = [
    ("label_sf",     "Spreading Factor",  str),
    ("label_mod",    "Modulation",        str),
    ("label_beacon", "Beacon ID",         str),
    ("label_pkt",    "Packet Type",       lambda x: PKT_LABELS.get(int(x), str(x))),
]
fig, axes = plt.subplots(1, 4, figsize=(14, 4))
for ax, (col, title, fmt) in zip(axes, TASKS):
    counts = df[col].value_counts().sort_index()
    labels = [fmt(k) for k in counts.index]
    bars = ax.bar(labels, counts.values, color="steelblue", edgecolor="white")
    ax.set_title(title)
    ax.set_xlabel(col)
    ax.set_ylabel("Count")
    for bar, v in zip(bars, counts.values):
        if v < 150:
            bar.set_color("tomato")
plt.tight_layout()
plt.savefig("eda5_class_balance.png", dpi=150)
\end{lstlisting}
Bars drawn in red indicate fewer than 150 samples; the 70/15/15 split will produce fewer than 30 test samples for that class, which is insufficient for reliable evaluation.

\subsection{Step 8 --- Operating-condition shift analysis}
Compare at least two features between the clean regime (C1, C2, and C4 single-station, DATA-only) and the contended regime (C5 cooperative-target ping-pong), entirely within the hardware dataset. The \texttt{regime} column added by the merge script identifies the source for each row.

\begin{enumerate}[leftmargin=2em]
	\item \textbf{\mbox{RSSI} shift:} Compute mean and standard deviation of \mbox{RSSI} for equivalent baseline conditions (SF7, +14~dBm, $\sim$3\,m), comparing C1's +14~dBm rows against the DATA rows of C5b. Half-duplex shadowing during ACK exchanges produces correlated dropouts that change \mbox{RSSI} variance even though the channel is physically the same.
	\item \textbf{Inter-arrival time shift:} Compare inter-arrival time histograms between the clean regime (tightly clustered $\sim$2\,s) and the contended regime (multi-modal: $\sim$1\,s DATA cadence plus ACK round-trip).
	\item For each comparison, state: the direction of the shift (e.g., ``contended-regime \mbox{RSSI} std is 1.5$\times$ the clean-regime value''), the likely physical cause (half-duplex shadowing, ACK collisions), and the implication for Module~4 (\emph{``a classifier trained only on the clean regime may misclassify ACK rows as DATA when deployed on contended traffic, leading to lower packet-type accuracy than the clean-only test set suggests''}).
\end{enumerate}

\begin{tcolorbox}[infobox={Why Operating-Condition Shift Matters for Module~4}]
	Module~4 will train classifiers on the hardware dataset collected here and evaluate them on held-out hardware data. The LH4-F experiment partitions the data along the \texttt{regime} axis: train on the clean regime (C1--C4), evaluate on the contended regime (C5), and measure the accuracy drop $\Delta = \mathrm{Acc}(\mathrm{clean}\!\to\!\mathrm{clean}) - \mathrm{Acc}(\mathrm{clean}\!\to\!\mathrm{contended})$. The distribution shift documented in this step provides the empirical explanation for that drop. Without this analysis, the Module~4 operating-condition shift experiment has no interpretive foundation.
\end{tcolorbox}

\subsection{Step 9 --- Data dictionary}
Write a \texttt{data\_dictionary.md} file documenting every column in \texttt{merged\_dataset.csv}:

\begin{tabularx}{\textwidth}{l l l X}
	\toprule
	\textbf{Column} & \textbf{Type} & \textbf{Units / Valid Range} & \textbf{Description} \\
	\midrule
	\texttt{campaign\_id} & string & C1, C2, C4, C5 & Experiment family identifier \\
	\texttt{run\_id} & string & unique per run & One \texttt{collect.py} invocation (ISO-timestamped) --- split-leakage key \\
	\texttt{regime} & string & ``clean''/``contended'' & Operating regime (clean = C1, C2, C4; contended = C5) \\
	\texttt{timestamp\_host} & ISO 8601 & UTC & Host reception time \\
	\texttt{rx\_timestamp\_ms} & uint32 & ms & Interceptor-side STM32 tick at RX \\
	\texttt{tx\_timestamp\_ms} & uint32 & ms & Transmitter-side STM32 tick (from packet payload) \\
	\texttt{inter\_arrival\_ms} & int32 & ms & RX timestamp delta between consecutive packets \\
	\texttt{beacon\_id} & uint8 & 1--255 & Sender Beacon~ID (\texttt{0x01}=Target~A, \texttt{0x02}=Target~B, \texttt{0xFF}=Threat) \\
	\texttt{seq\_num} & uint16 & $\geq 0$ & Per-station sequence counter (resets on firmware flash) \\
	\texttt{pkt\_type} & uint8 & 0=DATA, 1=ACK & Packet type from low 2 bits of \texttt{Modulation Code} byte \\
	\texttt{mod\_kind} & uint8 & 1=LoRa, 2=FSK, 3=OOK & Self-reported modulation kind (upper 6 bits of \texttt{Modulation Code} byte) \\
	\bottomrule
\end{tabularx}

\newpage

\begin{tabularx}{\textwidth}{l l l X}
	\toprule
	\textbf{Column} & \textbf{Type} & \textbf{Units / Valid Range} & \textbf{Description} \\
	\midrule
	\texttt{spread\_factor} & uint8 & 7--10 (LoRa); 0 (non-LoRa) & Self-reported SF (fixed by firmware compile-time define) \\
	\texttt{rssi\_dbm} & int8 & $-120$ to $0$~dBm & Received signal strength at interceptor \\
	\texttt{snr\_db} & float & $-20$ to $+15$~dB & SNR estimate (SX1276 \texttt{RegPktSnrValue} $\times$ 0.25) \\
	\texttt{freq\_error\_hz} & int16 & Hz & Crystal frequency offset between TX and interceptor \\
	\texttt{payload\_len} & uint8 & bytes & Radio payload length byte from packet \\
	\texttt{distance\_m} & float & 0.5--10~m & Measured Target--interceptor antenna separation \\
	\texttt{label\_sf} & uint8 & 7--10 & Ground-truth spreading factor (from campaign config) \\
	\texttt{label\_mod} & string & ``lora'' & Human-readable modulation label (alias of \texttt{mod\_kind\_label}) \\
	\texttt{mod\_kind\_label} & uint8 & 1=LoRa & Ground-truth modulation kind integer (from campaign config) \\
	\texttt{label\_beacon} & uint8 & 1 or 2 & Ground-truth Beacon~ID (sender identity, from packet) \\
	\texttt{label\_pkt} & uint8 & 0 or 1 & Ground-truth packet type (from packet) \\
	\texttt{split} & string & ``train''/``val''/``test'' & Dataset partition (stratified by \texttt{label\_sf}, keyed on \texttt{run\_id}) \\
	\bottomrule
\end{tabularx}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] \texttt{merged\_dataset.csv} contains $\geq$2{,}100 rows with the correct schema and a \texttt{split} column.
	\item[$\square$] No \texttt{[WARN]} lines in merge script output; \texttt{spread\_factor} matches \texttt{label\_sf} for all DATA rows.
	\item[$\square$] Both regimes well-represented: \texttt{regime} = \texttt{clean} and \texttt{regime} = \texttt{contended} both present with non-trivial row counts.
	\item[$\square$] All four classification tasks (SF, modulation, Beacon~ID, packet type) have at least 30 test samples per class.
	\item[$\square$] Train/val/test split is approximately 70/15/15\%.
	\item[$\square$] All five \mbox{EDA} plots generated and saved as \texttt{.png} files (Visualization 5 now covers all four classification tasks).
	\item[$\square$] Operating-condition shift analysis includes at least two quantitative comparisons (clean vs.\ contended mean/std or histogram divergence) entirely on hardware data.
	\item[$\square$] Data dictionary covers all columns with units and valid ranges, including the \texttt{pkt\_type} and \texttt{regime} columns.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 3-H Deliverable --- and Module 3 Final Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item \texttt{merged\_dataset.csv}: the complete labeled dataset with stratified split column.
		\item \texttt{merge\_dataset.py}: the merging script.
		\item Dataset summary: total rows, rows per campaign, rows per regime (clean / contended), rows per (Beacon~\mbox{ID}, \texttt{pkt\_type}) pair, rows per class for all four classification tasks (SF, modulation, Beacon~ID, packet type).
		\item All five \mbox{EDA} plots (\texttt{eda1\_} through \texttt{eda5\_}), each with a 2--3 sentence interpretive caption.
		\item Operating-condition shift analysis: at least two quantitative comparisons between the clean regime (C1, C2, and C4) and the contended regime (C5) with physical interpretation.
		\item \texttt{data\_dictionary.md}: complete column documentation.
		\item Known limitations section: indoor multipath effects, distance measurement uncertainty, half-duplex shadowing during Threat RX, any collection anomalies (power interruptions, firmware crashes, station desynchronization), and any campaigns that fell short of their target packet count.
		\item One-page Module~3 retrospective: what worked, what integration failures were most instructive, and what you would change before a second data collection campaign.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
