\documentclass[11pt,letterpaper]{article}
\newcommand{\labid}{4-G}
\newcommand{\labtitleshort}{Performance vs.\ Condition Analysis}
\input{LabHandout_Preamble}

\begin{document}
\labheader{4-G}{Stratified Accuracy by SNR, Distance, and Active-Station Count (with PAUSE)}

\section{Objective}
Produce the three stratified-accuracy curves that describe how each classifier \emph{degrades} (or does not) as operating conditions worsen: accuracy vs.\ \mbox{SNR}, accuracy vs.\ distance, and accuracy vs.\ active-station count (with PAUSE bursts present in the C5d/C5e sub-configurations). These curves are the robustness evidence used in the LH4-H final model selection rubric and they reveal which classifier family scales best to realistic deployment conditions.

\section{Prerequisites}
\begin{itemize}[leftmargin=2em]
	\item Lab Handouts~4-C through~4-F complete (RF, \mbox{SVM}, \mbox{MLP}, \mbox{CNN} all trained for each task).
	\item Campaign metadata present in \texttt{merged\_dataset.csv} (\texttt{distance\_m} column filled for C4 rows, \texttt{campaign\_id} C5a--C5e indicating active-station count and PAUSE-burst presence).
	\item \texttt{run\_id} column is populated (Module~3 LH3-F schema). The stratified-condition filters below select on \texttt{campaign\_id} for readability, but the underlying train/val/test split was keyed on \texttt{run\_id} (LH4-A Check~9), so no same-run leakage can occur even when a campaign family contains multiple runs.
\end{itemize}

\begin{tcolorbox}[infobox={Expected campaign\_id Values}]
This handout assumes the Module~3 canonical set \{C1, C2, C3, C4, C5a, C5b, C5c, C5d, C5e\}. If your \texttt{merged\_dataset.csv} uses different subcampaign labels, map them in \texttt{C5\_MAP} below \emph{and} update the C4 filter. The filters are cosmetic slicers, not leakage guards --- the leakage guard is the \texttt{run\_id}-keyed split from LH4-A.
\end{tcolorbox}

\section{Background: Average Accuracy Lies}
Two models can have identical top-line test accuracy while behaving very differently in the operating regime that matters. Model~A might do very well in high-\mbox{SNR} conditions and poorly in low-\mbox{SNR}; model~B might be uniformly mediocre. If the deployment environment is low-\mbox{SNR}-dominated, B is the better choice even though A has the higher headline number. Stratified accuracy surfaces this.

\section{Procedure}

\subsection{Step 1 --- Accuracy vs.\ SNR}
Bucket the test set by \texttt{snr} into five equal-count bins (quintiles). Compute each model's accuracy per bucket and plot as a line.

\begin{lstlisting}[language=Python, caption={\texttt{stratified\_accuracy.py} --- SNR-stratified evaluation.}]
import numpy as np, pandas as pd, joblib, torch
import matplotlib.pyplot as plt
from pathlib import Path

df = pd.read_csv("merged_dataset.csv")
df = df[df["obs_count"] > 1].reset_index(drop=True)
test = df[df["split"] == "test"].reset_index(drop=True)

def bucket_snr(test, n_buckets=5):
    edges = np.quantile(test["snr"], np.linspace(0, 1, n_buckets + 1))
    idx = np.digitize(test["snr"], edges[1:-1])
    centers = 0.5 * (edges[:-1] + edges[1:])
    return idx, centers, edges

def predict_rf(task, test_df):
    pre = joblib.load(f"artifacts/{task}/preprocessor.pkl")
    model = joblib.load(f"artifacts/{task}/rf.pkl")["model"]
    X = pre["scaler"].transform(
        test_df[pre["feature_cols"]].to_numpy(dtype=np.float32))
    y = pre["encoder"].transform(test_df[{
        "sf": "label_sf", "mod": "label_mod",
        "beacon": "label_beacon", "pkt": "label_pkt",
    }[task]])
    return model.predict(X), y

# Similar helpers for SVM, MLP, CNN. All return (y_pred, y_true).

def accuracy_by_bucket(idx, y_pred, y_true, n_buckets):
    accs = []
    for b in range(n_buckets):
        mask = (idx == b)
        if mask.sum() == 0:
            accs.append(np.nan); continue
        accs.append((y_pred[mask] == y_true[mask]).mean())
    return accs

task = "sf"   # example; repeat for each task
idx, centers, edges = bucket_snr(test)
plt.figure()
for model_name in ["rf", "svm", "mlp", "cnn"]:
    y_pred, y_true = predict(model_name, task, test)  # wrapper dispatcher
    accs = accuracy_by_bucket(idx, y_pred, y_true, len(centers))
    plt.plot(centers, accs, "-o", label=model_name.upper())
plt.xlabel("SNR (dB) bucket center")
plt.ylabel("Test accuracy")
plt.title(f"Accuracy vs. SNR --- task: {task}")
plt.legend(); plt.grid(True)
plt.tight_layout(); plt.savefig(f"figs/acc_vs_snr_{task}.png", dpi=150)
\end{lstlisting}

\begin{tcolorbox}[infobox={Sample Size per Bucket}]
	Quintiles give $\sim$20\% of the test set per bucket. With a $\sim$400-row test set, that is $\sim$80 samples per bucket. A single-bucket accuracy of 0.90 at $n=80$ has a 95\% Wilson CI of roughly $[0.82, 0.95]$. Be cautious about interpreting small between-model differences at individual bucket edges --- look for consistent patterns across buckets instead.
\end{tcolorbox}

\subsection{Step 2 --- Accuracy vs.\ distance}
Use Campaign~C4 (the distance sweep) exclusively for this plot. Distance is discrete in C4 (1, 3, 5, 8~m), so each distance is its own bucket rather than a quantile.

\begin{lstlisting}[language=Python]
c4 = df[df["campaign_id"] == "C4"].reset_index(drop=True)
distances = sorted(c4["distance_m"].unique())

for task in ["sf", "mod", "beacon", "pkt"]:
    plt.figure()
    for model_name in ["rf", "svm", "mlp", "cnn"]:
        y_pred, y_true = predict(model_name, task, c4)
        accs = [
            (y_pred[c4["distance_m"] == d] == y_true[c4["distance_m"] == d]).mean()
            for d in distances
        ]
        plt.plot(distances, accs, "-o", label=model_name.upper())
    plt.xlabel("Distance (m)")
    plt.ylabel("Test accuracy")
    plt.title(f"Accuracy vs. distance --- task: {task}")
    plt.legend(); plt.grid(True)
    plt.tight_layout()
    plt.savefig(f"figs/acc_vs_distance_{task}.png", dpi=150)
\end{lstlisting}

\subsection{Step 3 --- Accuracy vs.\ active-station count}
Use Campaign~C5. Each sub-configuration (C5a through C5e) corresponds to a different active-station count and PAUSE-burst pattern (C5a: Target~A only; C5b: A+B fixed ping-pong; C5c: A+B ALOHA; C5d: A+B + held-button PAUSE bursts; C5e: A+B ALOHA + intermittent PAUSE bursts). Plot accuracy as a function of station count, differentiating fixed vs.\ \mbox{ALOHA} scheduling and PAUSE-active vs.\ PAUSE-quiescent runs.

\begin{lstlisting}[language=Python]
# Map: (active stations, scheduling mode, threat PAUSE active?)
C5_MAP = {
    "C5a": (1, "fixed", False),
    "C5b": (2, "fixed", False),
    "C5c": (2, "aloha", False),
    "C5d": (3, "fixed", True),   # ping-pong + held-button PAUSE bursts
    "C5e": (3, "aloha", True),   # ALOHA + intermittent PAUSE bursts
}
c5 = df[df["campaign_id"].isin(C5_MAP)].reset_index(drop=True)

for task in ["sf", "mod", "beacon", "pkt"]:
    rows = []
    for cid, (n, mode, pause) in C5_MAP.items():
        sub = c5[c5["campaign_id"] == cid]
        if len(sub) == 0: continue
        for model_name in ["rf", "svm", "mlp", "cnn"]:
            y_pred, y_true = predict(model_name, task, sub)
            acc = (y_pred == y_true).mean()
            rows.append({"stations": n, "mode": mode, "pause": pause,
                         "model": model_name, "acc": acc})
    rdf = pd.DataFrame(rows)
    plt.figure()
    for (mode, pause, model_name), grp in rdf.groupby(["mode", "pause", "model"]):
        style  = "--" if mode == "aloha" else "-"
        marker = "x"  if pause             else "o"
        suffix = " +PAUSE" if pause else ""
        plt.plot(grp["stations"], grp["acc"], style + marker,
                 label=f"{model_name.upper()} ({mode}{suffix})")
    plt.xlabel("Active station count")
    plt.ylabel("Test accuracy")
    plt.title(f"Accuracy vs. active stations --- task: {task}")
    plt.legend(fontsize=7); plt.grid(True)
    plt.tight_layout()
    plt.savefig(f"figs/acc_vs_stations_{task}.png", dpi=150)
\end{lstlisting}

\begin{tcolorbox}[infobox={Note for the Packet-Type Task}]
The packet-type accuracy curve will look different from the other three tasks: C5a contains only DATA rows (the classifier has only one class to predict, so accuracy is mechanically 1.0), C5b/C5c add ACK rows, and C5d/C5e add PAUSE rows. Compare absolute numbers across sub-configurations cautiously --- the classifier's task difficulty changes with the active class set, not just with station count.
\end{tcolorbox}

\subsection{Step 4 --- Robustness score}
Define a simple, defensible robustness metric: the \emph{minimum} accuracy across all SNR buckets on the SF task. A model with high average accuracy but terrible worst-case accuracy is a worse deployment candidate than one with slightly lower average and higher worst-case.

\begin{lstlisting}[language=Python]
robustness = {}
idx, centers, _ = bucket_snr(test)
for model_name in ["rf", "svm", "mlp", "cnn"]:
    y_pred, y_true = predict(model_name, "sf", test)
    accs = accuracy_by_bucket(idx, y_pred, y_true, len(centers))
    robustness[model_name] = {
        "min":  float(np.nanmin(accs)),
        "mean": float(np.nanmean(accs)),
        "max":  float(np.nanmax(accs)),
    }
print(pd.DataFrame(robustness).T)
\end{lstlisting}

\subsection{Step 5 --- Narrative discussion (required for the deliverable)}
Write 2--3 paragraphs addressing:

\begin{enumerate}[leftmargin=2em]
	\item Which model degrades most gracefully as \mbox{SNR} drops? Is this the same model that has the highest average accuracy?
	\item Does the \mbox{CNN}'s temporal input give it an advantage in the multi-station regime (C5d, C5e), where DATA, ACK, and PAUSE packets from three different stations interleave unpredictably?
	\item Is there a model family whose accuracy curves become noticeably steeper (degrade faster) in the high-distance or high-contention regime? What does that suggest about how it would perform in a field deployment with obstructed line-of-sight?
\end{enumerate}

\begin{tcolorbox}[warning={Watch for Small-Bucket Artefacts}]
	If a distance or beacon-count bucket has $<$30 samples, do not plot a single accuracy point --- the variance is too high and the plot will mislead. Either merge it with an adjacent bucket, or annotate the point with an error bar computed from a Wilson CI. If you cannot meaningfully report a bucket, state that explicitly in the discussion.
\end{tcolorbox}

\section{Verification Checklist}
\begin{itemize}[leftmargin=2em]
	\item[$\square$] SNR-stratified accuracy plotted for all four tasks.
	\item[$\square$] Distance-stratified accuracy plotted (C4 data only).
	\item[$\square$] Active-station-count stratified accuracy plotted with fixed/\mbox{ALOHA} and PAUSE-active distinguished.
	\item[$\square$] Worst-case-bucket robustness table computed for at least the SF task.
	\item[$\square$] Narrative discussion (Step~5) written.
\end{itemize}

\begin{tcolorbox}[deliverable={Lab Handout 4-G Deliverable}]
	\begin{enumerate}[leftmargin=1.2em]
		\item Three stratified-accuracy figures per task (SNR, distance, station-count) = 12 figures total across the four classification tasks.
		\item Worst-case robustness table: min, mean, max bucket-accuracy per model on the SF task.
		\item A ``robustness ranking'' per task: order the four models (RF, SVM, MLP, CNN) from most to least robust, with one sentence justifying each position.
		\item 2--3 paragraph narrative discussion addressing the three questions in Step~5.
	\end{enumerate}
\end{tcolorbox}

\printhandoutglossary

\end{document}
