\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}
\begin{document}
	
	\begin{center}
		{\Large\textbf{EE800/820 Weekly Status Report}}
	\end{center}
	
	\vspace{0.5em}
	
	\noindent\textbf{Student:} Jude Eschete\\
	\noindent\textbf{Project Advisor:} Bernard Yett\\
	\textbf{Week of:} February 16, 2026\\
	\textbf{Project:} An Educational Framework for AI-Driven RF Signal Processing on FPGA
	
	\vspace{1em}
	\hrule
	\vspace{1em}
	
	\section*{Accomplishments This Week}
	
	\begin{itemize}[noitemsep]
		\item Received all ordered hardware components, including the STM32 Nucleo-L476RG microcontrollers and RF transceiver modules.
		\item Completed development of the "Engagement Simulator," a Python-based software tool to model threat/target engagement scenarios and visualize binary data streams before hardware deployment.
		\item Implemented core simulation constants and utility helpers for signal sequence generation (jamming, telemetry, status updates).
		\item Refined the project directory structure to support the simulation application alongside the hardware codebase.
	\end{itemize}
	
	\section*{Goals for Next Week}
	
	\begin{itemize}[noitemsep]
		\item Validate engagement scenarios using the completed Simulator to generate baseline data.
		\item Begin hardware bring-up: Flash STM32 boards and verify basic SPI communication with the RF modules.
		\item Finalize the "Module 1" lesson plan document, incorporating details from the specific hardware received.
	\end{itemize}
	
	\section*{Potential Blockers}
	
	\begin{itemize}[noitemsep]
		\item Ensuring simulation data accurately reflects hardware behavior during the transition to physical implementation.
	\end{itemize}
	
\end{document}